---
title: "Protected KBA Indicator Comparison"
format:
  html:
    self-contained: true
    toc: true
    toc_float: true
    code_folding: hide
    theme: cosmo
    math: "mathjax"
params:
  indicator_a: 5731
  indicator_b: 3963
execute:
  echo: false
  warning: false
  message: false
date: today
---

## Comparing Indicators `r params$indicator_a` and `r params$indicator_b`

<!--
This report is an exploratory comparison between two CEPALSTAT indicators that are supposed to
represent the same underlying concept, following the same methodology (as documented in their
technical sheets / metadata, both pulled live from the CEPALSTAT API below). In practice their
published values diverge, and the goal of this notebook is to characterize *how* and *where* they
diverge -- by country, by year, and structurally (source, methodology, dimensions) -- so that a
root cause can be identified and written up in the Summary of Findings section at the end.
-->

```{r setup}
library(tidyverse)
library(magrittr)
library(readxl)
library(glue)
library(knitr)
library(gt)
library(reactable)
library(plotly)
library(scales)
library(here)

source(here("Scripts/utils.R"))

id_a <- params$indicator_a
id_b <- params$indicator_b

DIM_COUNTRY <- 208
DIM_YEAR    <- 29117

# Strip stray <br> tags that sometimes leak into CEPALSTAT's published metadata text, and collapse
# extra whitespace, so long-form fields (definition, methodology, comments) read cleanly in the report.
strip_html <- function(x) {
  x %>% str_remove_all("<br\\s*/?>") %>% str_squish()
}
```

## Technical Sheet Comparison

*This section pulls each indicator's live metadata ("technical sheet") from the CEPALSTAT API and
compares them field by field. Since both indicators are supposed to follow the same methodology,
any differences here (definition, calculation methodology, data source, dimensions) are the first
place to look for an explanation of diverging values.*

```{r fetch metadata}
meta_a <- get_indicator_metadata(id_a) %>% mutate(value = strip_html(value))
meta_b <- get_indicator_metadata(id_b) %>% mutate(value = strip_html(value))

name_a <- meta_a %>% filter(variable == "indicator_name") %>% pull(value)
name_b <- meta_b %>% filter(variable == "indicator_name") %>% pull(value)

label_a <- glue("Indicator {id_a}: {name_a}")
label_b <- glue("Indicator {id_b}: {name_b}")
```

**`r label_a`**

**`r label_b`**

### Key Fields

The fields below drive the actual calculation and are the most likely source of any methodological
divergence. A "Match" / "Differs" flag is shown for each.

```{r display key metadata fields, results='asis'}
key_fields <- c("indicator_name", "definition", "calculation_methodology", "comments",
                "unit_measurement", "periodicity", "geographic_coverage")
key_fields <- intersect(key_fields, union(meta_a$variable, meta_b$variable))

get_field <- function(meta, field) {
  val <- meta$value[meta$variable == field]
  if (length(val) == 0 || is.na(val) || val == "") "*(not provided)*" else val
}

for (f in key_fields) {
  val_a <- get_field(meta_a, f)
  val_b <- get_field(meta_b, f)
  differs <- !identical(val_a, val_b)

  field_title <- str_to_title(str_replace_all(f, "_", " "))
  flag <- if (differs) "&#9888;&#65039; **Differs**" else "&#9989; Match"

  cat(glue("#### {field_title} -- {flag}\n\n"))
  cat(glue("**Indicator {id_a}:**\n\n{val_a}\n\n"))
  cat(glue("**Indicator {id_b}:**\n\n{val_b}\n\n"))
  cat("\n---\n\n")
}
```

### Data Source

```{r compare source}
source_a <- get_indicator_source(id_a) %>% slice(1)
source_b <- get_indicator_source(id_b) %>% slice(1)

tibble(
  Indicator = c(as.character(id_a), as.character(id_b)),
  Source    = c(
    glue("{source_a$organization_acronym} / {source_a$description}"),
    glue("{source_b$organization_acronym} / {source_b$description}")
  )
) %>%
  gt() %>%
  tab_header(title = "Assigned Data Source")
```

### Dimensions

*If one indicator carries an extra dimension the other doesn't (e.g. a breakdown the other
collapses or omits), that's a structural difference worth flagging on its own.*

```{r compare dimensions}
dims_a <- get_indicator_dimensions(id_a)
dims_b <- get_indicator_dimensions(id_b)

only_dims_a <- setdiff(dims_a$name, dims_b$name)
only_dims_b <- setdiff(dims_b$name, dims_a$name)
```

`r if (length(only_dims_a) > 0) glue("Dimensions present **only** in indicator {id_a}: {paste(only_dims_a, collapse = ', ')}.") else glue("No dimensions unique to indicator {id_a}.")`

`r if (length(only_dims_b) > 0) glue("Dimensions present **only** in indicator {id_b}: {paste(only_dims_b, collapse = ', ')}.") else glue("No dimensions unique to indicator {id_b}.")`

### Full Metadata Field Listing (Appendix)

*Every metadata field returned by the API for either indicator, side by side, in case something
outside the key fields above explains the discrepancy.*

```{r full metadata appendix}
meta_compare <- full_join(
  meta_a %>% rename(value_a = value),
  meta_b %>% rename(value_b = value),
  by = "variable"
) %>%
  mutate(differs = coalesce(value_a, "") != coalesce(value_b, "")) %>%
  arrange(desc(differs), variable)

reactable(
  meta_compare %>%
    rename(!!as.character(glue("Indicator {id_a}")) := value_a, !!as.character(glue("Indicator {id_b}")) := value_b),
  columns = list(
    variable = colDef(name = "Field"),
    differs  = colDef(name = "Differs?")
  ),
  defaultPageSize = 8,
  searchable = TRUE,
  highlight = TRUE,
  bordered = TRUE,
  wrap = TRUE,
  theme = reactableTheme(
    style = list(fontFamily = "Segoe UI, Roboto, Helvetica, sans-serif", fontSize = "14px")
  )
)
```

## Data Comparison

*This section pulls the actual published data for both indicators and compares them on a
country-by-country basis.*

```{r fetch data}
raw_a <- get_cepalstat_data(id_a)
raw_b <- get_cepalstat_data(id_b)

base_dim_cols <- paste0("dim_", c(DIM_COUNTRY, DIM_YEAR))

extra_dims_a <- setdiff(raw_a %>% select(starts_with("dim_")) %>% names(), base_dim_cols)
extra_dims_b <- setdiff(raw_b %>% select(starts_with("dim_")) %>% names(), base_dim_cols)
```

`r if (length(extra_dims_a) > 0) glue("Indicator {id_a} carries additional dimensions beyond country/year ({paste(extra_dims_a, collapse = ', ')}); these are summed to a single country/year value below.") else ""`

`r if (length(extra_dims_b) > 0) glue("Indicator {id_b} carries additional dimensions beyond country/year ({paste(extra_dims_b, collapse = ', ')}); these are summed to a single country/year value below.") else ""`

```{r summarize and label}
country_dim <- get_dimension_table(DIM_COUNTRY) %>% transmute(dim_208 = as.character(id), country = name)
year_dim    <- get_dimension_table(DIM_YEAR)    %>% transmute(dim_29117 = as.character(id), year = name)

# Collapse to a single country/year value (summing over any extra dimensions), then attach labels.
summarize_indicator <- function(raw) {
  raw %>%
    mutate(
      value = as.numeric(value),
      dim_208 = as.character(dim_208),
      dim_29117 = as.character(dim_29117)
    ) %>%
    group_by(dim_208, dim_29117) %>%
    summarize(value = sum_or_na(value), .groups = "drop") %>%
    left_join(country_dim, by = "dim_208") %>%
    left_join(year_dim, by = "dim_29117") %>%
    mutate(year_num = as.numeric(year))
}

sum_a <- summarize_indicator(raw_a)
sum_b <- summarize_indicator(raw_b)
```

### Year Coverage

*Before comparing values, it's worth checking whether the two indicators even cover the same
years -- if they don't, any country-level comparison has to fall back on a proxy year for
whichever indicator is missing the other's most recent year.*

```{r year coverage}
year_min_a <- min(sum_a$year_num, na.rm = TRUE)
year_max_a <- max(sum_a$year_num, na.rm = TRUE)
year_min_b <- min(sum_b$year_num, na.rm = TRUE)
year_max_b <- max(sum_b$year_num, na.rm = TRUE)

# The year used for each indicator in the country-level comparison below: its own most recent
# available year. When the two indicators' year ranges don't overlap, this amounts to a
# nearest-available-year proxy comparison rather than a true same-year comparison (see note below).
compare_year_a <- year_max_a
compare_year_b <- year_max_b

range_label <- function(lo, hi) if (lo == hi) as.character(lo) else glue("{lo}-{hi}")

tibble(
  Indicator = c(as.character(id_a), as.character(id_b)),
  `Year Range` = c(range_label(year_min_a, year_max_a), range_label(year_min_b, year_max_b)),
  `Year Used for Comparison` = c(as.character(compare_year_a), as.character(compare_year_b))
) %>%
  gt() %>%
  tab_header(title = "Year Coverage by Indicator")
```

`r if (compare_year_a == compare_year_b) glue("Both indicators share year **{compare_year_a}**, so the comparison below is a true same-year comparison.") else glue("**Indicator {id_a}** only has data for **{range_label(year_min_a, year_max_a)}**, while **indicator {id_b}** covers **{range_label(year_min_b, year_max_b)}**. Since the two indicators don't share a common year, the country-level comparison below uses each indicator's own most recent available year as the closest practical cross-check: **{compare_year_a}** for indicator {id_a} against **{compare_year_b}** for indicator {id_b}. This is not a true same-year comparison -- part of any difference below could reflect indicator {id_b}'s own year-over-year change from {compare_year_b} onward (see the Country-Level Time Series below) rather than a genuine methodological discrepancy between the two indicators.")`

### Country Coverage

*It's also worth checking whether the two indicators cover the same set of countries -- a coverage
mismatch (e.g. one including a "World" or regional aggregate the other doesn't) can itself explain
part of the apparent divergence.*

```{r country coverage diagnostic}
countries_a <- sum_a %>% distinct(country) %>% pull()
countries_b <- sum_b %>% distinct(country) %>% pull()

only_in_a <- setdiff(countries_a, countries_b)
only_in_b <- setdiff(countries_b, countries_a)
```

`r if (length(only_in_a) > 0) glue("Entities reported **only** by indicator {id_a}: {paste(only_in_a, collapse = ', ')}.") else glue("No entities unique to indicator {id_a}.")`

`r if (length(only_in_b) > 0) glue("Entities reported **only** by indicator {id_b}: {paste(only_in_b, collapse = ', ')}.") else glue("No entities unique to indicator {id_b}.")`

```{r restrict to lac and build comparison df}
lac_countries <- iso %>% pull(std_name)

# Country-level comparison at each indicator's own comparison year (see Year Coverage above) --
# not a country/year join, since the two indicators' year ranges don't overlap.
comp <- full_join(
  sum_a %>% filter(country %in% lac_countries, year_num == compare_year_a) %>% select(country, value_a = value),
  sum_b %>% filter(country %in% lac_countries, year_num == compare_year_b) %>% select(country, value_b = value),
  by = "country"
) %>%
  mutate(
    abs_diff = abs(value_a - value_b),
    perc_diff = case_when(
      is.na(value_a) | is.na(value_b) ~ NA_real_,
      value_a == value_b ~ 0,
      value_b == 0 ~ NA_real_,
      TRUE ~ round(abs_diff / abs(value_b) * 100, 1)
    ),
    status = case_when(
      is.na(value_a) & is.na(value_b) ~ "Missing in Both",
      is.na(value_a) ~ sprintf("Only in Indicator %s", id_b),
      is.na(value_b) ~ sprintf("Only in Indicator %s", id_a),
      TRUE ~ "Present in Both"
    ),
    flag_large_diff = !is.na(perc_diff) & perc_diff > 20
  )
```

### Data Availability by Country and Year

*This shows each indicator's full published history (not just the single comparison year used
below) -- useful for seeing the overall shape of the year mismatch described above at a glance.*

```{r coverage heatmap prep}
full_status <- full_join(
  sum_a %>% filter(country %in% lac_countries) %>% select(country, year, year_num, value_a = value),
  sum_b %>% filter(country %in% lac_countries) %>% select(country, year, year_num, value_b = value),
  by = c("country", "year", "year_num")
) %>%
  mutate(status = case_when(
    is.na(value_a) & is.na(value_b) ~ "Missing in Both",
    is.na(value_a) ~ sprintf("Only in Indicator %s", id_b),
    is.na(value_b) ~ sprintf("Only in Indicator %s", id_a),
    TRUE ~ "Present in Both"
  ))

all_combinations <- expand_grid(
  country = sort(unique(full_status$country)),
  year_num = min(full_status$year_num, na.rm = TRUE):max(full_status$year_num, na.rm = TRUE)
) %>%
  mutate(year = as.character(year_num))

coverage_check <- all_combinations %>%
  left_join(full_status %>% select(country, year, status), by = c("country", "year")) %>%
  mutate(status = ifelse(is.na(status), "Missing in Both", status))
```

```{r coverage heatmap}
status_levels <- c("Present in Both", sprintf("Only in Indicator %s", id_a),
                    sprintf("Only in Indicator %s", id_b), "Missing in Both")
status_colors <- setNames(c("#e6e6e6", "#99ccff", "#ffcc99", "white"), status_levels)

p <- coverage_check %>%
  mutate(
    country = factor(country, levels = sort(unique(country), decreasing = TRUE)),
    status = factor(status, levels = status_levels)
  ) %>%
  ggplot(aes(x = year, y = country, fill = status,
             text = paste("Country:", country, "<br>Year:", year, "<br>Status:", status))) +
  geom_tile(color = "white") +
  scale_fill_manual(values = status_colors) +
  labs(title = "Data Availability by Country and Year", x = NULL, y = NULL, fill = "Coverage") +
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    legend.position = "bottom",
    legend.title = element_text(face = "bold")
  )

ggplotly(p, tooltip = "text") %>%
  layout(legend = list(orientation = "h", y = -0.2))
```

### Value Comparison

*Comparing each country's indicator `r id_a` value in `r compare_year_a` against its indicator
`r id_b` value in `r compare_year_b` -- the closest available cross-check given the year mismatch
noted above. The dashed line in the scatter plot marks perfect agreement; points far from it are
where the two indicators disagree most.*

```{r summary statistics}
summary_stats <- comp %>%
  filter(status == "Present in Both") %>%
  summarise(
    "Countries compared" = n(),
    "Mean % difference" = round(mean(perc_diff[is.finite(perc_diff)], na.rm = TRUE), 1),
    "Median % difference" = round(median(perc_diff[is.finite(perc_diff)], na.rm = TRUE), 1),
    "% of countries differing by >20%" = round(mean(flag_large_diff, na.rm = TRUE) * 100, 1),
    "Mean absolute difference" = round(mean(abs_diff[is.finite(abs_diff)], na.rm = TRUE), 2)
  )

tibble(Metric = names(summary_stats), Value = unlist(summary_stats)) %>%
  mutate(Value = case_when(
    grepl("%", Metric) ~ paste0(Value, "%"),
    TRUE ~ as.character(Value)
  )) %>%
  gt() %>%
  tab_header(
    title = "Summary of Differences",
    subtitle = glue("Indicator {id_a} ({compare_year_a}) vs indicator {id_b} ({compare_year_b})")
  )
```

```{r scatter comparison}
p <- comp %>%
  filter(status == "Present in Both") %>%
  mutate(flag_large_diff = factor(flag_large_diff, levels = c(FALSE, TRUE),
                                   labels = c("Within ±20%", "Above ±20%"))) %>%
  ggplot(aes(
    x = value_a, y = value_b, color = flag_large_diff,
    text = paste("Country:", country,
                 "<br>Indicator", id_a, glue("({compare_year_a}):"), round(value_a, 2),
                 "<br>Indicator", id_b, glue("({compare_year_b}):"), round(value_b, 2),
                 "<br>% Diff:", perc_diff)
  )) +
  geom_point(alpha = 0.8, size = 1.4) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray60") +
  scale_color_manual(values = c("Within ±20%" = "#99ccff", "Above ±20%" = "#cc6666"), guide = "none") +
  labs(
    title = "Value Comparison Between Indicators",
    subtitle = glue("Indicator {id_a} ({compare_year_a}) vs indicator {id_b} ({compare_year_b})"),
    x = glue("Indicator {id_a} ({compare_year_a})"), y = glue("Indicator {id_b} ({compare_year_b})")
  ) +
  theme_minimal() +
  theme(plot.title = element_text(face = "bold", size = 14), plot.subtitle = element_text(size = 11))

ggplotly(p, tooltip = "text")
```

### Country-Level Time Series

*Indicator `r id_b`'s full published history plotted against indicator `r id_a`'s single available
year, country by country. Use this to judge whether indicator `r id_a`'s value looks like a
plausible continuation of indicator `r id_b`'s trend, or a sharp break from it.*

```{r country time series, fig.height=16, warning=FALSE}
ts_data <- bind_rows(
  sum_a %>% filter(country %in% lac_countries) %>% mutate(indicator = as.character(id_a)) %>%
    select(country, year_num, value, indicator),
  sum_b %>% filter(country %in% lac_countries) %>% mutate(indicator = as.character(id_b)) %>%
    select(country, year_num, value, indicator)
) %>%
  filter(!is.na(value))

p <- ts_data %>%
  ggplot(aes(
    x = year_num, y = value, color = indicator, group = indicator,
    text = paste("Country:", country, "<br>Year:", year_num,
                 "<br>Indicator:", indicator, "<br>Value:", round(value, 2))
  )) +
  geom_line(linewidth = 0.7, alpha = 0.8) +
  geom_point(size = 0.9) +
  facet_wrap(~ country, scales = "free_y", ncol = 3) +
  scale_color_manual(values = setNames(c("#99ccff", "#cc6666"), c(as.character(id_a), as.character(id_b)))) +
  labs(title = "Indicator Values by Country", x = "Year", y = "Value", color = "Indicator") +
  theme_minimal() +
  theme(plot.title = element_text(face = "bold", size = 14), legend.position = "bottom")

ggplotly(p, tooltip = "text") %>% layout(legend = list(orientation = "h", y = -0.05))
```

### Which Countries Diverge Most

```{r country level pct diff}
p <- comp %>%
  filter(status == "Present in Both") %>%
  arrange(perc_diff) %>%
  mutate(country = factor(country, levels = country)) %>%
  ggplot(aes(x = perc_diff, y = country,
             text = paste("Country:", country, "<br>% Diff:", perc_diff))) +
  geom_col(fill = "#99ccff") +
  labs(
    title = "Percent Difference by Country",
    subtitle = glue("Indicator {id_a} ({compare_year_a}) vs indicator {id_b} ({compare_year_b})"),
    x = "Percent Difference", y = NULL
  ) +
  theme_minimal() +
  theme(plot.title = element_text(face = "bold", size = 14), plot.subtitle = element_text(size = 11))

ggplotly(p, tooltip = "text")
```

### Most Divergent Countries

```{r divergent table}
reactable(
  comp %>%
    filter(status == "Present in Both") %>%
    mutate(value_a = round(value_a, 2), value_b = round(value_b, 2), abs_diff = round(abs_diff, 2)) %>%
    arrange(desc(abs(perc_diff))) %>%
    select(country, value_a, value_b, abs_diff, perc_diff) %>%
    rename(
      !!as.character(glue("Indicator {id_a} ({compare_year_a})")) := value_a,
      !!as.character(glue("Indicator {id_b} ({compare_year_b})")) := value_b
    ),
  defaultPageSize = 10,
  searchable = TRUE,
  highlight = TRUE,
  bordered = TRUE,
  theme = reactableTheme(
    style = list(fontFamily = "Segoe UI, Roboto, Helvetica, sans-serif", fontSize = "14px")
  )
)
```

## Summary of Findings

```{r auto summary text, results='asis'}
n_both <- comp %>% filter(status == "Present in Both") %>% nrow()
pct_large <- summary_stats[["% of countries differing by >20%"]]
mean_pct <- summary_stats[["Mean % difference"]]

fields_differ <- length(key_fields[sapply(key_fields, function(f) !identical(get_field(meta_a, f), get_field(meta_b, f)))]) > 0

year_sentence <- if (compare_year_a == compare_year_b) {
  glue("Both indicators share a common year ({compare_year_a}), so this is a true same-year comparison.")
} else {
  glue("Indicators {id_a} and {id_b} do not share a common year -- this comparison uses indicator {id_a}'s {compare_year_a} values against indicator {id_b}'s {compare_year_b} values as the closest available proxy, so some of the gap below could simply reflect one additional year of change rather than a true methodological discrepancy.")
}

coverage_sentence <- if (length(only_in_a) > 0 || length(only_in_b) > 0) {
  "Country coverage also differs between the two indicators (see the Country Coverage section above), which by itself accounts for some of the apparent mismatch."
} else {
  "Both indicators cover the same set of countries, so coverage is not driving the divergence."
}

fields_sentence <- if (fields_differ) {
  "One or more key technical sheet fields differ between the indicators (flagged above) -- this is the most likely place to start when explaining the gap."
} else {
  "The key technical sheet fields (definition, methodology, source, comments) read as identical, so the divergence is likely coming from the underlying data itself rather than a documented methodological change."
}

cat(glue(
  "Across **{n_both}** countries compared, the mean absolute percent difference is **{mean_pct}%**, ",
  "and **{pct_large}%** of countries differ by more than 20%. ",
  "{year_sentence} {coverage_sentence} {fields_sentence}"
))
```

<!--
Use this space to record the analyst's own conclusions after reviewing the visuals above. Consider:

- **Theoretical / methodological**: Do the technical sheets actually describe the same calculation?
  Any differences in scope, formula, or definitions found above?
- **Data source**: Are the two indicators pulling from different underlying datasets or vintages
  (see Data Source section)? Could one be using an older/newer release of the same source?
- **Coverage**: Do the two indicators cover the same countries (see Country Coverage)? Note the two
  indicators don't share a common year (see Year Coverage) -- part of any gap could simply reflect
  the newer indicator's own year-over-year change rather than a genuine discrepancy.
- **Pattern of divergence**: Is the gap concentrated in specific countries (see Which Countries
  Diverge Most), or does the single-year indicator's value look like a plausible continuation of
  the other indicator's historical trend (see Country-Level Time Series)?
-->

*(Add manual conclusions here once the visuals above have been reviewed.)*

**Theoretical / methodological:**

**Data source:**

**Coverage / data-related:**
