---
title: "Protected KBA Indicator Comparison"
format:
  html:
    self-contained: true
    toc: true
    toc_float: true
    code_folding: hide
    theme: cosmo
    math: "mathjax"
params:
  indicator_a: 5731
  indicator_b: 3963
execute:
  echo: false
  warning: false
  message: false
date: today
---

## Comparing Indicators `r params$indicator_a` and `r params$indicator_b`

<!--
This report is an exploratory comparison between two CEPALSTAT indicators that are supposed to
represent the same underlying concept, following the same methodology (as documented in their
technical sheets / metadata, both pulled live from the CEPALSTAT API below). In practice their
published values diverge, and the goal of this notebook is to characterize *how* and *where* they
diverge -- by country, by year, and structurally (source, methodology, dimensions) -- so that a
root cause can be identified and written up in the Summary of Findings section at the end.
-->

```{r setup}
library(tidyverse)
library(magrittr)
library(readxl)
library(glue)
library(knitr)
library(gt)
library(reactable)
library(plotly)
library(scales)
library(here)

source(here("Scripts/utils.R"))

id_a <- params$indicator_a
id_b <- params$indicator_b

DIM_COUNTRY <- 208
DIM_YEAR    <- 29117

# Strip stray <br> tags that sometimes leak into CEPALSTAT's published metadata text, and collapse
# extra whitespace, so long-form fields (definition, methodology, comments) read cleanly in the report.
strip_html <- function(x) {
  x %>% str_remove_all("<br\\s*/?>") %>% str_squish()
}
```

## Technical Sheet Comparison

*This section pulls each indicator's live metadata ("technical sheet") from the CEPALSTAT API and
compares them field by field. Since both indicators are supposed to follow the same methodology,
any differences here (definition, calculation methodology, data source, dimensions) are the first
place to look for an explanation of diverging values.*

```{r fetch metadata}
meta_a <- get_indicator_metadata(id_a) %>% mutate(value = strip_html(value))
meta_b <- get_indicator_metadata(id_b) %>% mutate(value = strip_html(value))

name_a <- meta_a %>% filter(variable == "indicator_name") %>% pull(value)
name_b <- meta_b %>% filter(variable == "indicator_name") %>% pull(value)

label_a <- glue("Indicator {id_a}: {name_a}")
label_b <- glue("Indicator {id_b}: {name_b}")
```

**`r label_a`**

**`r label_b`**

### Key Fields

The fields below drive the actual calculation and are the most likely source of any methodological
divergence. A "Match" / "Differs" flag is shown for each.

```{r display key metadata fields, results='asis'}
key_fields <- c("indicator_name", "definition", "calculation_methodology", "comments",
                "unit_measurement", "periodicity", "geographic_coverage")
key_fields <- intersect(key_fields, union(meta_a$variable, meta_b$variable))

get_field <- function(meta, field) {
  val <- meta$value[meta$variable == field]
  if (length(val) == 0 || is.na(val) || val == "") "*(not provided)*" else val
}

for (f in key_fields) {
  val_a <- get_field(meta_a, f)
  val_b <- get_field(meta_b, f)
  differs <- !identical(val_a, val_b)

  field_title <- str_to_title(str_replace_all(f, "_", " "))
  flag <- if (differs) "&#9888;&#65039; **Differs**" else "&#9989; Match"

  cat(glue("#### {field_title} -- {flag}\n\n"))
  cat(glue("**Indicator {id_a}:**\n\n{val_a}\n\n"))
  cat(glue("**Indicator {id_b}:**\n\n{val_b}\n\n"))
  cat("\n---\n\n")
}
```

### Data Source

```{r compare source}
source_a <- get_indicator_source(id_a) %>% slice(1)
source_b <- get_indicator_source(id_b) %>% slice(1)

tibble(
  Indicator = c(as.character(id_a), as.character(id_b)),
  Source    = c(
    glue("{source_a$organization_acronym} / {source_a$description}"),
    glue("{source_b$organization_acronym} / {source_b$description}")
  )
) %>%
  gt() %>%
  tab_header(title = "Assigned Data Source")
```

### Dimensions

*If one indicator carries an extra dimension the other doesn't (e.g. a breakdown the other
collapses or omits), that's a structural difference worth flagging on its own.*

```{r compare dimensions}
dims_a <- get_indicator_dimensions(id_a)
dims_b <- get_indicator_dimensions(id_b)

only_dims_a <- setdiff(dims_a$name, dims_b$name)
only_dims_b <- setdiff(dims_b$name, dims_a$name)
```

`r if (length(only_dims_a) > 0) glue("Dimensions present **only** in indicator {id_a}: {paste(only_dims_a, collapse = ', ')}.") else glue("No dimensions unique to indicator {id_a}.")`

`r if (length(only_dims_b) > 0) glue("Dimensions present **only** in indicator {id_b}: {paste(only_dims_b, collapse = ', ')}.") else glue("No dimensions unique to indicator {id_b}.")`

### Full Metadata Field Listing (Appendix)

*Every metadata field returned by the API for either indicator, side by side, in case something
outside the key fields above explains the discrepancy.*

```{r full metadata appendix}
meta_compare <- full_join(
  meta_a %>% rename(value_a = value),
  meta_b %>% rename(value_b = value),
  by = "variable"
) %>%
  mutate(differs = coalesce(value_a, "") != coalesce(value_b, "")) %>%
  arrange(desc(differs), variable)

reactable(
  meta_compare %>%
    rename(!!as.character(glue("Indicator {id_a}")) := value_a, !!as.character(glue("Indicator {id_b}")) := value_b),
  columns = list(
    variable = colDef(name = "Field"),
    differs  = colDef(name = "Differs?")
  ),
  defaultPageSize = 8,
  searchable = TRUE,
  highlight = TRUE,
  bordered = TRUE,
  wrap = TRUE,
  theme = reactableTheme(
    style = list(fontFamily = "Segoe UI, Roboto, Helvetica, sans-serif", fontSize = "14px")
  )
)
```

## Data Comparison

*This section pulls the actual published data for both indicators and compares them on a
country-by-country, year-by-year basis.*

```{r fetch data}
raw_a <- get_cepalstat_data(id_a)
raw_b <- get_cepalstat_data(id_b)

base_dim_cols <- paste0("dim_", c(DIM_COUNTRY, DIM_YEAR))

extra_dims_a <- setdiff(raw_a %>% select(starts_with("dim_")) %>% names(), base_dim_cols)
extra_dims_b <- setdiff(raw_b %>% select(starts_with("dim_")) %>% names(), base_dim_cols)
```

`r if (length(extra_dims_a) > 0) glue("Indicator {id_a} carries additional dimensions beyond country/year ({paste(extra_dims_a, collapse = ', ')}); these are summed to a single country/year value below.") else ""`

`r if (length(extra_dims_b) > 0) glue("Indicator {id_b} carries additional dimensions beyond country/year ({paste(extra_dims_b, collapse = ', ')}); these are summed to a single country/year value below.") else ""`

```{r summarize and label}
country_dim <- get_dimension_table(DIM_COUNTRY) %>% transmute(dim_208 = as.character(id), country = name)
year_dim    <- get_dimension_table(DIM_YEAR)    %>% transmute(dim_29117 = as.character(id), year = name)

# Collapse to a single country/year value (summing over any extra dimensions), then attach labels.
summarize_indicator <- function(raw) {
  raw %>%
    mutate(
      value = as.numeric(value),
      dim_208 = as.character(dim_208),
      dim_29117 = as.character(dim_29117)
    ) %>%
    group_by(dim_208, dim_29117) %>%
    summarize(value = sum_or_na(value), .groups = "drop") %>%
    left_join(country_dim, by = "dim_208") %>%
    left_join(year_dim, by = "dim_29117")
}

sum_a <- summarize_indicator(raw_a)
sum_b <- summarize_indicator(raw_b)
```

### Country Coverage

*Before comparing values, it's worth checking whether the two indicators even cover the same set
of countries -- a coverage mismatch (e.g. one including a "World" or regional aggregate the other
doesn't) can itself explain part of the apparent divergence.*

```{r country coverage diagnostic}
countries_a <- sum_a %>% distinct(country) %>% pull()
countries_b <- sum_b %>% distinct(country) %>% pull()

only_in_a <- setdiff(countries_a, countries_b)
only_in_b <- setdiff(countries_b, countries_a)
```

`r if (length(only_in_a) > 0) glue("Entities reported **only** by indicator {id_a}: {paste(only_in_a, collapse = ', ')}.") else glue("No entities unique to indicator {id_a}.")`

`r if (length(only_in_b) > 0) glue("Entities reported **only** by indicator {id_b}: {paste(only_in_b, collapse = ', ')}.") else glue("No entities unique to indicator {id_b}.")`

```{r restrict to lac and build comparison df}
lac_countries <- iso %>% pull(std_name)

comp <- full_join(
  sum_a %>% filter(country %in% lac_countries) %>% select(country, year, value_a = value),
  sum_b %>% filter(country %in% lac_countries) %>% select(country, year, value_b = value),
  by = c("country", "year")
) %>%
  mutate(
    year_num = as.numeric(year),
    abs_diff = abs(value_a - value_b),
    perc_diff = case_when(
      is.na(value_a) | is.na(value_b) ~ NA_real_,
      value_a == value_b ~ 0,
      value_b == 0 ~ NA_real_,
      TRUE ~ round(abs_diff / abs(value_b) * 100, 1)
    ),
    status = case_when(
      is.na(value_a) & is.na(value_b) ~ "Missing in Both",
      is.na(value_a) ~ sprintf("Only in Indicator %s", id_b),
      is.na(value_b) ~ sprintf("Only in Indicator %s", id_a),
      TRUE ~ "Present in Both"
    ),
    flag_large_diff = !is.na(perc_diff) & perc_diff > 20
  )
```

### Data Availability by Country and Year

```{r coverage heatmap prep}
all_combinations <- expand_grid(
  country = sort(unique(comp$country)),
  year = as.character(min(comp$year_num, na.rm = TRUE):max(comp$year_num, na.rm = TRUE))
)

coverage_check <- all_combinations %>%
  left_join(comp %>% select(country, year, status), by = c("country", "year")) %>%
  mutate(status = ifelse(is.na(status), "Missing in Both", status))
```

```{r coverage heatmap}
status_levels <- c("Present in Both", sprintf("Only in Indicator %s", id_a),
                    sprintf("Only in Indicator %s", id_b), "Missing in Both")
status_colors <- setNames(c("#e6e6e6", "#99ccff", "#ffcc99", "white"), status_levels)

p <- coverage_check %>%
  mutate(
    country = factor(country, levels = sort(unique(country), decreasing = TRUE)),
    status = factor(status, levels = status_levels)
  ) %>%
  ggplot(aes(x = year, y = country, fill = status,
             text = paste("Country:", country, "<br>Year:", year, "<br>Status:", status))) +
  geom_tile(color = "white") +
  scale_fill_manual(values = status_colors) +
  labs(title = "Data Availability by Country and Year", x = NULL, y = NULL, fill = "Coverage") +
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    legend.position = "bottom",
    legend.title = element_text(face = "bold")
  )

ggplotly(p, tooltip = "text") %>%
  layout(legend = list(orientation = "h", y = -0.2))
```

### Value Comparison

*Comparing values only where both indicators report a country/year. The dashed line marks perfect
agreement -- points far from it are where the two indicators disagree most.*

```{r summary statistics}
summary_stats <- comp %>%
  filter(status == "Present in Both") %>%
  summarise(
    "Country-years compared" = n(),
    "Mean % difference" = round(mean(perc_diff[is.finite(perc_diff)], na.rm = TRUE), 1),
    "Median % difference" = round(median(perc_diff[is.finite(perc_diff)], na.rm = TRUE), 1),
    "% of pairs differing by >20%" = round(mean(flag_large_diff, na.rm = TRUE) * 100, 1),
    "Mean absolute difference" = round(mean(abs_diff[is.finite(abs_diff)], na.rm = TRUE), 2)
  )

tibble(Metric = names(summary_stats), Value = unlist(summary_stats)) %>%
  mutate(Value = case_when(
    grepl("%", Metric) ~ paste0(Value, "%"),
    TRUE ~ as.character(Value)
  )) %>%
  gt() %>%
  tab_header(title = "Summary of Differences", subtitle = "Country-years present in both indicators")
```

```{r scatter comparison}
p <- comp %>%
  filter(status == "Present in Both") %>%
  mutate(flag_large_diff = factor(flag_large_diff, levels = c(FALSE, TRUE),
                                   labels = c("Within ±20%", "Above ±20%"))) %>%
  ggplot(aes(
    x = value_a, y = value_b, color = flag_large_diff,
    text = paste("Country:", country, "<br>Year:", year,
                 "<br>Indicator", id_a, ":", round(value_a, 2),
                 "<br>Indicator", id_b, ":", round(value_b, 2),
                 "<br>% Diff:", perc_diff)
  )) +
  geom_point(alpha = 0.8, size = 1.4) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray60") +
  scale_color_manual(values = c("Within ±20%" = "#99ccff", "Above ±20%" = "#cc6666"), guide = "none") +
  labs(
    title = "Value Comparison Between Indicators",
    x = glue("Indicator {id_a}"), y = glue("Indicator {id_b}")
  ) +
  theme_minimal() +
  theme(plot.title = element_text(face = "bold", size = 14))

ggplotly(p, tooltip = "text")
```

```{r percent diff over time}
p <- comp %>%
  filter(status == "Present in Both") %>%
  ggplot(aes(
    x = year_num, y = perc_diff,
    text = paste("Country:", country, "<br>Year:", year, "<br>% Diff:", perc_diff)
  )) +
  geom_jitter(aes(color = flag_large_diff), width = 0.2, height = 0, size = 1.4, alpha = 0.8) +
  scale_color_manual(values = c("FALSE" = "#99ccff", "TRUE" = "#cc6666"), guide = "none") +
  labs(
    title = "Percent Difference Over Time",
    subtitle = "Red points differ by more than 20% between the two indicators",
    x = "Year", y = "Percent Difference"
  ) +
  theme_minimal() +
  theme(plot.title = element_text(face = "bold", size = 14), plot.subtitle = element_text(size = 11))

ggplotly(p, tooltip = "text")
```

### Country-Level Time Series

*Direct comparison of both indicators' values, country by country, over time. Use this to spot
whether divergence is a constant offset, a growing gap, or isolated to specific years.*

```{r country time series, fig.height=16, warning=FALSE}
ts_data <- comp %>%
  filter(!(is.na(value_a) & is.na(value_b))) %>%
  select(country, year_num, value_a, value_b) %>%
  pivot_longer(cols = c(value_a, value_b), names_to = "indicator", values_to = "value") %>%
  mutate(indicator = recode(indicator, value_a = as.character(id_a), value_b = as.character(id_b))) %>%
  filter(!is.na(value))

p <- ts_data %>%
  ggplot(aes(
    x = year_num, y = value, color = indicator, group = indicator,
    text = paste("Country:", country, "<br>Year:", year_num,
                 "<br>Indicator:", indicator, "<br>Value:", round(value, 2))
  )) +
  geom_line(linewidth = 0.7, alpha = 0.8) +
  geom_point(size = 0.9) +
  facet_wrap(~ country, scales = "free_y", ncol = 3) +
  scale_color_manual(values = setNames(c("#99ccff", "#cc6666"), c(as.character(id_a), as.character(id_b)))) +
  labs(title = "Indicator Values by Country", x = "Year", y = "Value", color = "Indicator") +
  theme_minimal() +
  theme(plot.title = element_text(face = "bold", size = 14), legend.position = "bottom")

ggplotly(p, tooltip = "text") %>% layout(legend = list(orientation = "h", y = -0.05))
```

### Which Countries and Years Diverge Most

```{r country level pct diff}
country_diff <- comp %>%
  filter(status == "Present in Both") %>%
  group_by(country) %>%
  summarise(mean_perc_diff = round(mean(perc_diff, na.rm = TRUE), 1),
            n_years = n(), .groups = "drop") %>%
  arrange(desc(mean_perc_diff))

p <- country_diff %>%
  mutate(country = factor(country, levels = rev(country))) %>%
  ggplot(aes(x = mean_perc_diff, y = country,
             text = paste("Country:", country, "<br>Mean % Diff:", mean_perc_diff, "<br>N years:", n_years))) +
  geom_col(fill = "#99ccff") +
  labs(title = "Mean Percent Difference by Country", x = "Mean % Difference", y = NULL) +
  theme_minimal() +
  theme(plot.title = element_text(face = "bold", size = 14))

ggplotly(p, tooltip = "text")
```

```{r year level pct diff}
year_diff <- comp %>%
  filter(status == "Present in Both") %>%
  group_by(year_num) %>%
  summarise(mean_perc_diff = round(mean(perc_diff, na.rm = TRUE), 1), n_countries = n(), .groups = "drop")

p <- year_diff %>%
  ggplot(aes(x = year_num, y = mean_perc_diff,
             text = paste("Year:", year_num, "<br>Mean % Diff:", mean_perc_diff, "<br>N countries:", n_countries))) +
  geom_line(color = "#99ccff", linewidth = 0.9) +
  geom_point(color = "#3377bb", size = 1.6) +
  labs(title = "Mean Percent Difference by Year",
       subtitle = "A rising trend suggests the two indicators are drifting apart over time (e.g. differing vintages/updates)",
       x = "Year", y = "Mean % Difference") +
  theme_minimal() +
  theme(plot.title = element_text(face = "bold", size = 14), plot.subtitle = element_text(size = 11))

ggplotly(p, tooltip = "text")
```

### Most Divergent Country-Years

```{r divergent table}
reactable(
  comp %>%
    filter(status == "Present in Both") %>%
    mutate(value_a = round(value_a, 2), value_b = round(value_b, 2), abs_diff = round(abs_diff, 2)) %>%
    arrange(desc(abs(perc_diff))) %>%
    select(country, year, value_a, value_b, abs_diff, perc_diff) %>%
    rename(!!as.character(glue("Indicator {id_a}")) := value_a, !!as.character(glue("Indicator {id_b}")) := value_b),
  defaultPageSize = 10,
  searchable = TRUE,
  highlight = TRUE,
  bordered = TRUE,
  theme = reactableTheme(
    style = list(fontFamily = "Segoe UI, Roboto, Helvetica, sans-serif", fontSize = "14px")
  )
)
```

## Summary of Findings

```{r auto summary text, results='asis'}
n_both <- comp %>% filter(status == "Present in Both") %>% nrow()
pct_large <- summary_stats[["% of pairs differing by >20%"]]
mean_pct <- summary_stats[["Mean % difference"]]

fields_differ <- length(key_fields[sapply(key_fields, function(f) !identical(get_field(meta_a, f), get_field(meta_b, f)))]) > 0

coverage_sentence <- if (length(only_in_a) > 0 || length(only_in_b) > 0) {
  "Coverage also differs between the two indicators (see the Country Coverage section above), which by itself accounts for some of the apparent mismatch."
} else {
  "Both indicators cover the same set of entities, so coverage is not driving the divergence."
}

fields_sentence <- if (fields_differ) {
  "One or more key technical sheet fields differ between the indicators (flagged above) -- this is the most likely place to start when explaining the gap."
} else {
  "The key technical sheet fields (definition, methodology, source, comments) read as identical, so the divergence is likely coming from the underlying data itself rather than a documented methodological change."
}

cat(glue(
  "Across **{n_both}** country-year pairs where both indicators report a value, the mean absolute ",
  "percent difference is **{mean_pct}%**, and **{pct_large}%** of pairs differ by more than 20%. ",
  "{coverage_sentence} {fields_sentence}"
))
```

<!--
Use this space to record the analyst's own conclusions after reviewing the visuals above. Consider:

- **Theoretical / methodological**: Do the technical sheets actually describe the same calculation?
  Any differences in scope, formula, or definitions found above?
- **Data source**: Are the two indicators pulling from different underlying datasets or vintages
  (see Data Source section)? Could one be using an older/newer release of the same source?
- **Coverage**: Do the two indicators cover the same countries and years (see Country Coverage)?
- **Pattern of divergence**: Is the gap a constant offset, growing over time, or concentrated in
  specific countries (see Country-Level Time Series and the by-country/by-year charts)?
-->

*(Add manual conclusions here once the visuals above have been reviewed.)*

**Theoretical / methodological:**

**Data source:**

**Coverage / data-related:**
