---
title: "CEPALSTAT QC Dashboard"
format:
  dashboard:
    theme: cosmo
    scrolling: true
server: shiny
---

<!--
Interactive port of Scripts/qc_report.qmd (the original QC report) into a Quarto Dashboard
with a Shiny backend. Same checks, same flags (±20% difference threshold, z-score outlier
threshold of 3), same underlying logic - just reactive instead of static, and driven by
sidebar filters (Country, Year, and one per extra dimension) instead of facet grids. Because
the filters let you narrow down to what you actually want to look at, this doesn't need the
"facet by country" or "first two dimensions only" versions of each chart - an Inspect Country
dropdown replaces those.

Everything server-side lives in `context: server` chunks; Quarto assembles them into one
server function in document order regardless of where they sit in the file. They're kept
next to the page/card markup they power (rather than one giant chunk up top) so each section
of this file is self-contained to read and edit.

`context: server` chunks are extracted for the server function, not executed during the
top-to-bottom pass that knits the UI - so any package whose functions are called directly in
UI chunks (reactableOutput(), plotlyOutput()) has to be loaded outside of context: server, or
it won't exist yet when the UI chunk referencing it gets knitted.
-->

```{r}
# Loaded in a plain (non-server) chunk on purpose - see note above.
library(reactable)
library(plotly)
```

```{r}
#| context: server
# ---- Setup: libraries, palette, and data-loading functions ----

library(shiny)
library(plotly)
library(reactable)
library(ggplot2)
library(dplyr)
library(tidyr)
library(tibble)
library(glue)
library(scales)
library(readxl)
library(magrittr)
library(stringr)
library(lubridate)
library(here)
library(fs)

source(here("Scripts/utils.R")) # also loads the global `meta` and `iso` tables

pal <- list(
  nav = "#17485A", header = "#3F8EAB", highlight = "#20B2B0", background = "#C7DDE4",
  before = "#2a78d6", after = "#eb6834", 
  flag = "#d03b3b", good = "#0ca30c",
  neutral = "#c3c2b7", grid = "#e1e0d9", fill_both = "#e1e0d9", fill_missing = "#fcfcfb",
  ink = "#0b0b0b", ink2 = "#52514e",
  categorical = c("#C53532", "#2E4655", "#5C9EA9", "#D27F66", "#83B99B", "#B59B93", "#C77D1B",  "#707173", "#759A82", "#617783", "#BBC6CF")
)

theme_qc <- function() {
  theme_minimal(base_size = 12) %+replace%
    theme(
      plot.title = element_text(face = "bold", size = rel(1.1), color = pal$ink, hjust = 0),
      axis.text = element_text(color = pal$ink2),
      axis.title = element_text(color = pal$ink2),
      panel.grid.major = element_line(color = pal$grid, linewidth = 0.4),
      panel.grid.minor = element_blank(),
      legend.title = element_text(face = "bold", color = pal$ink),
      legend.text = element_text(color = pal$ink2)
    )
}

# No title/subtitle here on purpose: the Quarto Dashboard card already shows a title bar
# (via each chunk's #| title: option), so a second, plotly-internal title was redundant - and
# the fixed top margin added to make room for it was the main source of the extra whitespace.
style_plotly <- function(gp, legend_bottom = TRUE) {
  if (legend_bottom) {
    gp <- gp %>% layout(legend = list(orientation = "h", y = -0.15))
  } else {
    gp <- gp %>% layout(showlegend = FALSE)
  }
  gp
}

# A blank plotly canvas with a centered message - used instead of req()-blocking an output
# outright, so a card always shows something (the real chart, or a clear reason it's empty)
# rather than going silently blank with no explanation.
empty_plotly <- function(msg) {
  plotly_empty(type = "scatter", mode = "markers") %>%
    layout(
      xaxis = list(visible = FALSE), yaxis = list(visible = FALSE),
      annotations = list(list(text = msg, showarrow = FALSE, font = list(size = 14, color = pal$ink2)))
    )
}

# Pixel height for a facet_wrap grid of country panels (3 columns), sized to the number of
# countries currently in view. Used with plotlyOutput(height=) via a renderUI wrapper, since
# that's the reliable way to size a Shiny-rendered plotly widget (unlike the fig.height chunk
# option, which only applies to a chunk's own auto-printed output, not renderPlotly()).
facet_height_px <- function(n_facets, ncol = 3, per_row = 170, base = 100) {
  paste0(base + ceiling(max(n_facets, 1) / ncol) * per_row, "px")
}

# Blends a hex color toward light gray, for a "duller/muted" version of the same hue - used on
# the Published Preview page (solid full color = current internal value, muted dot-dash =
# currently published value), so a before/after pair never needs two hand-picked colors.
mute_color <- function(hex, amount = 0.55) {
  rgb_val <- grDevices::col2rgb(hex) / 255
  gray_target <- 0.75
  muted <- rgb_val * (1 - amount) + gray_target * amount
  grDevices::rgb(muted[1], muted[2], muted[3])
}

# Wraps a tall plotlyOutput in a fixed-height, scrollable container. Keeps the card's own
# height "regular" (not growing to match a 30+ country facet grid) while the plot itself still
# renders at its full, correctly-proportioned height inside - capping the container's height
# directly (instead of this wrapper) would just squish the panels again.
scrollable_plot <- function(output_id, plot_height_px, max_visible_px = 700) {
  div(
    style = glue("max-height:{max_visible_px}px; overflow-y:auto;"),
    plotlyOutput(output_id, height = plot_height_px)
  )
}

callout <- function(text, type = c("note", "good", "flag")) {
  type <- match.arg(type)
  color <- switch(type, note = pal$before, good = pal$good, flag = pal$flag)
  icon  <- switch(type, note = "ℹ", good = "✓", flag = "⚠")
  glue(
    '<div style="border-left:4px solid {color}; background:#f7f7f5; padding:0.7rem 1rem; ',
    'border-radius:4px; margin:0.5rem 0;"><strong>{icon}</strong>&nbsp; {text}</div>'
  )
}

# Cached per indicator_id - the expensive part (read files, resolve dimension names via API).
.qc_raw_cache <- new.env(parent = emptyenv())

load_qc_raw <- function(indicator_id, use_cache = TRUE) {
  cache_key <- as.character(indicator_id)
  if (use_cache && exists(cache_key, envir = .qc_raw_cache)) {
    return(get(cache_key, envir = .qc_raw_cache))
  }

  comp_path <- here(glue("Data/Checks/comp_id{indicator_id}.xlsx"))
  if (!file_exists(comp_path)) {
    stop(glue(
      "No comparison file found for indicator {indicator_id} at Data/Checks/comp_id{indicator_id}.xlsx. ",
      "Run the indicator through the processing pipeline (export = TRUE) first."
    ))
  }

  comp_detailed <- read_xlsx(comp_path)
  comp_detailed %<>%
    mutate(across(.cols = matches("dim"), .fns = as.character)) %>%
    mutate(across(.cols = matches("value|^abs_diff$|^perc_diff$"), .fns = as.numeric)) %>%
    mutate(across(.cols = starts_with("flag_"), .fns = as.logical))

  indicator_name <- meta %>% filter(id == indicator_id) %>% pull(indicator)
  if (length(indicator_name) == 0) indicator_name <- NA_character_
  source_vals <- meta %>% filter(id == indicator_id) %>% pull(source)
  source_val <- if (length(source_vals) == 0) NA_character_ else source_vals

  files <- dir_ls(here("Data/Cleaned"), regexp = paste0("id", indicator_id, "_\\d{4}-\\d{2}-\\d{2}T\\d{6}"))
  data_date <- if (length(files) > 0) {
    latest_file <- tibble(
      file = files,
      datetime = str_extract(files, "\\d{4}-\\d{2}-\\d{2}T\\d{6}") %>% ymd_hms(truncated = 1)
    ) %>% arrange(desc(datetime)) %>% slice(1) %>% pull(file)
    str_extract(latest_file, "\\d{4}-\\d{2}-\\d{2}T\\d{6}")
  } else {
    NA_character_
  }

  ind_dims <- comp_detailed %>% select(ends_with("_label")) %>% names()
  base_dims <- c("dim_29117_label", "dim_208_label") # year, country
  extra_dims <- setdiff(ind_dims, base_dims)
  extra_dim_ids <- str_extract(extra_dims, "\\d+")

  extra_dim_names <- if (length(extra_dim_ids) > 0) {
    get_indicator_dimensions(indicator_id) %>% filter(id %in% extra_dim_ids) %>% pull(name)
  } else {
    character(0)
  }

  has_public_data <- any(!is.na(comp_detailed$value_pub))
  has_extra_dims <- length(extra_dims) >= 1
  dim_labels <- setNames(extra_dim_names, extra_dims)
  dim_members <- setNames(
    lapply(extra_dims, function(dc) sort(unique(na.omit(comp_detailed[[dc]])))),
    extra_dims
  )

  all_countries <- sort(unique(comp_detailed$dim_208_label))
  all_years <- sort(unique(as.numeric(comp_detailed$dim_29117_label)))

  result <- list(
    indicator_id = indicator_id, indicator_name = indicator_name,
    source = source_val, data_date = data_date,
    comp_detailed = comp_detailed, all_countries = all_countries,
    year_min = min(all_years), year_max = max(all_years),
    extra_dims = extra_dims, dim_labels = dim_labels, dim_members = dim_members,
    has_public_data = has_public_data, has_extra_dims = has_extra_dims,
    n_records = nrow(comp_detailed)
  )
  if (use_cache) assign(cache_key, result, envir = .qc_raw_cache)
  result
}

# Cheap - filters already-loaded data by country/year/dimension-member selections, then
# re-aggregates to country/year and recomputes every flag. Always operates on the FULL data -
# every country, year, and dimension member - per diff_threshold, adjustable from the sidebar's
# Threshold control (the original report's ±20% and z-score-of-3 conventions are the defaults).
# The sidebar's Highlight controls (Country, Year, dimension members) are for spotlighting a
# slice on top of this, not for narrowing what's included - see the Country control below and
# the Dimension Analysis page for where that highlighting is currently wired in.
build_filtered_view <- function(raw, diff_threshold = 20, outlier_z = 3) {
  comp_detailed <- raw$comp_detailed

  comp <- comp_detailed %>%
    group_by(dim_208, dim_29117, dim_208_label, dim_29117_label) %>%
    summarize(
      value_data = if (all(is.na(value_data))) NA_real_ else sum(value_data, na.rm = TRUE),
      value_pub  = if (all(is.na(value_pub)))  NA_real_ else sum(value_pub,  na.rm = TRUE),
      .groups = "drop"
    ) %>%
    mutate(abs_diff = abs(value_data - value_pub),
           perc_diff = round((abs_diff / value_pub) * 100, 2)) %>%
    mutate(
      flag_large_diff = perc_diff > diff_threshold,
      flag_missing_entry = is.na(value_data),
      flag_new_entry = is.na(value_pub),
      flag_some_na = is.na(value_data) | is.na(value_pub)
    ) %>%
    mutate(status = case_when(
      is.na(value_data) & is.na(value_pub) ~ "Missing in Both",
      is.na(value_data) ~ "Old Only",
      is.na(value_pub) ~ "New Only",
      TRUE ~ "Present in Both"
    )) %>%
    group_by(dim_208_label) %>%
    mutate(
      z_within_country = round(as.numeric(scale(value_data)), 1),
      outlier_within_country = abs(z_within_country) > outlier_z
    ) %>%
    ungroup() %>%
    mutate(flag_some_value = ifelse(outlier_within_country == TRUE | flag_large_diff == TRUE, TRUE, FALSE))

  all_combinations <- expand_grid(
    country = raw$all_countries,
    year = as.character(raw$year_min:raw$year_max)
  )
  coverage_check <- all_combinations %>%
    left_join(comp, by = c("country" = "dim_208_label", "year" = "dim_29117_label")) %>%
    mutate(status = ifelse(is.na(status), "Missing in Both", status))

  countries_with_flags <- comp %>%
    filter(flag_some_na == FALSE & flag_some_value == TRUE) %>%
    distinct(dim_208_label) %>%
    pull()

  list(
    comp = comp, comp_detailed = comp_detailed, coverage_check = coverage_check,
    countries = raw$all_countries, countries_with_flags = countries_with_flags
  )
}

# Finds member values of one dimension (e.g. a specific year, country, or extra-dimension
# category) that are ENTIRELY new or entirely missing relative to the public data - i.e. the
# member never appears with a value on one side at all, not just a few missing rows. Used for
# the Overview page's "added"/"removed" KPIs. Labels each item with dim_label for display when
# results from multiple dimensions get combined (e.g. "Year: 2025", "Country: Haiti").
find_dim_changes <- function(df, dim_col, dim_label) {
  member_status <- df %>%
    group_by(.data[[dim_col]]) %>%
    summarise(
      has_data = any(!is.na(value_data)),
      has_pub  = any(!is.na(value_pub)),
      .groups = "drop"
    )
  added <- member_status %>% filter(has_data, !has_pub) %>% pull(.data[[dim_col]])
  removed <- member_status %>% filter(has_pub, !has_data) %>% pull(.data[[dim_col]])
  list(
    added = if (length(added) > 0) paste0(dim_label, ": ", added) else character(0),
    removed = if (length(removed) > 0) paste0(dim_label, ": ", removed) else character(0)
  )
}

# Condenses find_dim_changes()'s one-row-per-member output (e.g. "Year: 2020", "Year: 2021",
# "Economic activity: Construction") down to one row per DIMENSION, with its members
# comma-joined (e.g. "Year: 2020, 2021", "Economic activity: Construction") - for display only;
# the un-condensed count of individual members is used for the tile's headline number.
condense_changes <- function(items) {
  if (length(items) == 0) return(character(0))
  dim_label <- sub(":.*$", "", items)
  member <- sub("^[^:]*: ", "", items)
  tibble(dim_label = dim_label, member = member) %>%
    group_by(dim_label) %>%
    summarise(joined = paste(member, collapse = ", "), .groups = "drop") %>%
    mutate(line = paste0(dim_label, ": ", joined)) %>%
    pull(line)
}

# Aggregates one highlighted country's values by year for the Published Preview page - one row
# per year x dimension member (or a single "Value" pseudo-member when the indicator has no
# extra dimension), mirroring how CEPALSTAT's own published chart aggregates a country/
# dimension selection down to a single series.
build_published_preview <- function(raw, country, dim_selections = list()) {
  df <- raw$comp_detailed %>% filter(dim_208_label == country)

  if (raw$has_extra_dims) {
    dim_col <- raw$extra_dims[1]
    sel <- dim_selections[[dim_col]]
    # "__ALL__" is the dimension filter's "(All)" sentinel (see dim_filters_ui) - not a real
    # member value, so it means "don't narrow" rather than "match nothing."
    if (!is.null(sel) && length(sel) > 0 && !("__ALL__" %in% sel)) {
      df <- df %>% filter(.data[[dim_col]] %in% sel)
    }
    df <- df %>% rename(member = all_of(dim_col))
  } else {
    df <- df %>% mutate(member = "Value")
  }

  df %>%
    group_by(dim_29117_label, member) %>%
    summarize(
      value_data = if (all(is.na(value_data))) NA_real_ else sum(value_data, na.rm = TRUE),
      value_pub  = if (all(is.na(value_pub)))  NA_real_ else sum(value_pub,  na.rm = TRUE),
      .groups = "drop"
    ) %>%
    mutate(year = as.numeric(dim_29117_label)) %>%
    select(-dim_29117_label)
}
```

```{r}
#| context: server
# ---- Reactives ----

qc_data <- eventReactive(input$load, {
  req(input$indicator_id)
  tryCatch(
    load_qc_raw(input$indicator_id),
    error = function(e) validate(need(FALSE, conditionMessage(e)))
  )
}, ignoreNULL = FALSE)

# Always the full dataset (every country/year/dimension member), aggregated at the
# sidebar's Threshold setting - see build_filtered_view()'s comment for why this no longer
# takes any country/year/dimension-member narrowing.
filtered_data <- reactive({
  d <- qc_data()
  req(d, input$threshold)
  build_filtered_view(d, diff_threshold = input$threshold)
})
```

```{r}
#| context: server
# ---- Sidebar filter UI: rebuilt whenever a new indicator loads ----

output$status_ui <- renderUI({
  d <- qc_data()
  name_display <- if (is.na(d$indicator_name)) glue("Indicator {d$indicator_id}") else d$indicator_name
  div(
    style = glue("padding:0.65rem 0.8rem; background:{pal$background}; border-radius:8px;"),
    div(style = glue("font-weight:700; font-size:1.05em; color:{pal$nav}; line-height:1.3;"), name_display),
    div(
      style = glue("margin-top:0.3rem; font-size:0.78em; color:{pal$header};"),
      glue("Vintage {if (is.na(d$data_date)) 'unknown' else d$data_date}")
    )
  )
})

# Year range and dimension-member selections here are for spotlighting a slice on the charts
# (not narrowing the data - see build_filtered_view()). Currently wired into the Published
# Preview page (year zooms its x-axis, dimension members each become their own line there) -
# not yet connected to any other chart.
output$year_slider_ui <- renderUI({
  d <- qc_data()
  sliderInput("year_range", "Years", min = d$year_min, max = d$year_max,
              value = c(d$year_min, d$year_max), step = 1, sep = "")
})

output$dim_filters_ui <- renderUI({
  d <- qc_data()
  if (!d$has_extra_dims) return(NULL)
  tagList(lapply(d$extra_dims, function(dc) {
    members <- d$dim_members[[dc]]
    choices <- c("(All)" = "__ALL__", setNames(members, members))
    selectizeInput(
      inputId = paste0("dimfilter_", dc),
      label = d$dim_labels[[dc]],
      choices = choices,
      selected = "__ALL__",
      multiple = TRUE
    )
  }))
})

# The Highlight section's Country control - defaults to Latin America and the Caribbean, and
# drives the single-country charts (Inspect Country's old job) across every page.
output$highlight_country_ui <- renderUI({
  d <- qc_data()
  default_country <- if ("Latin America and the Caribbean" %in% d$all_countries) {
    "Latin America and the Caribbean"
  } else {
    d$all_countries[1]
  }
  selectInput("country", "Country", choices = d$all_countries, selected = default_country)
})
```

## {.sidebar}

```{r}
numericInput("indicator_id", "Indicator ID", value = 4243, min = 1, step = 1)
```

```{r}
actionButton("load", "Load indicator", class = "btn-primary w-100")
```

---

```{r}
uiOutput("status_ui")
```

---

**Highlight**

```{r}
uiOutput("highlight_country_ui")
```

```{r}
uiOutput("year_slider_ui")
```

```{r}
uiOutput("dim_filters_ui")
```

---

**Threshold**

```{r}
sliderInput("threshold", "Difference threshold (%)", min = 0, max = 100, value = 20, step = 1)
```

# Overview

```{r}
#| context: server
output$summary_table <- renderReactable({
  d <- qc_data()
  info <- list(
    "Indicator ID" = d$indicator_id,
    "Indicator Name" = d$indicator_name,
    "Data Vintage" = d$data_date,
    "Data Source" = d$source,
    "Dimension Names" = if (length(d$dim_labels) > 0) paste(d$dim_labels, collapse = ", ") else "(none)",
    "Time Coverage" = paste(d$year_min, "-", d$year_max),
    "Number of Countries" = length(d$all_countries),
    "Total Records" = d$n_records,
    "Has Published Data" = ifelse(d$has_public_data, "Yes", "No (new indicator)")
  )
  df <- tibble(Metric = names(info), Value = as.character(unlist(info)))
  reactable(df, columns = list(Metric = colDef(width = 220)), bordered = TRUE, highlight = TRUE)
})

# Color a percentage by severity: green <10%, yellow 10-30%, red >30% (also used for the
# large-difference and mean-%-difference tiles, per the same scale).
severity_color <- function(pct) {
  if (is.na(pct)) return(pal$neutral)
  if (pct < 10) pal$good else if (pct <= 30) "#eda100" else pal$flag
}

# Small hand-built line icons (thin stroke, no fill) matching the reference gallery's outlined
# icon style - plain inline SVG rather than an icon font/package, so there's no new dependency
# and the stroke color always matches the tile exactly.
svg_icon <- function(name, color) {
  attrs <- glue('viewBox="0 0 24 24" width="22" height="22" fill="none" stroke="{color}" stroke-width="2.2" stroke-linecap="round" stroke-linejoin="round"')
  body <- switch(name,
    warning = glue('<path d="M12 3 L22 20 L2 20 Z"/><line x1="12" y1="9" x2="12" y2="14"/><circle cx="12" cy="17" r="1" fill="{color}" stroke="none"/>'),
    delta   = '<path d="M12 4 L21 19 L3 19 Z"/>',
    plus    = '<line x1="12" y1="5" x2="12" y2="19"/><line x1="5" y1="12" x2="19" y2="12"/>',
    minus   = '<line x1="5" y1="12" x2="19" y2="12"/>'
  )
  HTML(glue('<svg {attrs}>{body}</svg>'))
}

# A stat-card KPI tile in the style of a typical gallery dashboard: a circular outlined icon
# badge on the left, and to its right a small bold label, the big number, and a small caption
# below (one line per item, e.g. for a list of added/removed dimension members) - all in the
# one theme color, on a solid light tint of it rather than a faint wash.
kpi_tile <- function(icon, value, label, color, caption = NULL) {
  icon_badge <- div(
    style = glue(
      "flex:0 0 auto; width:46px; height:46px; border-radius:50%; border:2.5px solid {color}; ",
      "display:flex; align-items:center; justify-content:center;"
    ),
    icon
  )
  text_col <- div(
    style = "flex:1 1 auto; min-width:0;",
    div(style = glue("font-size:0.78rem; font-weight:700; color:{color};"), label),
    div(style = glue("font-size:1.8rem; font-weight:700; color:{color}; line-height:1.15; margin-top:0.05rem;"), value),
    if (!is.null(caption) && length(caption) > 0) {
      div(
        style = glue("font-size:0.7rem; color:{color}; margin-top:0.25rem; line-height:1.4;"),
        lapply(caption, function(x) tagList(x, br()))
      )
    }
  )
  div(
    style = glue(
      "flex:1 1 0; min-width:190px; background:{color}22; border-radius:10px; ",
      "padding:0.75rem 0.9rem; display:flex; align-items:center; gap:0.75rem;"
    ),
    icon_badge, text_col
  )
}

# Four KPI tiles: how much changed (large-difference % and mean % difference, both computed on
# the full data at the sidebar's Threshold), and what changed structurally (dimension members -
# year, country, or the extra dimension - that are entirely new or entirely gone relative to
# the public data, not just a few missing rows).
output$kpi_boxes_ui <- renderUI({
  d <- qc_data()

  if (d$has_public_data) {
    fd <- filtered_data()
    present <- fd$comp %>% filter(status == "Present in Both")
    pct_large <- round(mean(present$flag_large_diff, na.rm = TRUE) * 100, 1)
    mean_diff <- round(mean(present$perc_diff[is.finite(present$perc_diff)], na.rm = TRUE), 1)

    dim_specs <- list(c("dim_29117_label", "Year"), c("dim_208_label", "Country"))
    if (d$has_extra_dims) {
      dim_col <- d$extra_dims[1]
      dim_specs <- c(dim_specs, list(c(dim_col, d$dim_labels[[dim_col]])))
    }
    changes <- lapply(dim_specs, function(spec) find_dim_changes(d$comp_detailed, spec[1], spec[2]))
    added <- unlist(lapply(changes, `[[`, "added"))
    removed <- unlist(lapply(changes, `[[`, "removed"))

    pct_large_display <- if (is.na(pct_large)) "N/A" else glue("{pct_large}%")
    mean_diff_display <- if (is.na(mean_diff)) "N/A" else glue("{mean_diff}%")
    large_color <- severity_color(pct_large)
    mean_color <- severity_color(mean_diff)
  } else {
    pct_large_display <- "N/A"
    mean_diff_display <- "N/A"
    added <- character(0)
    removed <- character(0)
    large_color <- pal$neutral
    mean_color <- pal$neutral
  }

  added_color <- if (length(added) > 0) pal$good else pal$neutral
  removed_color <- if (length(removed) > 0) pal$flag else pal$neutral

  div(
    style = "display:flex; gap:0.6rem; flex-wrap:wrap;",
    kpi_tile(
      svg_icon("warning", large_color), pct_large_display, "Large Differences", large_color,
      caption = glue("of values differ by more than {input$threshold}% from what's published")
    ),
    kpi_tile(
      svg_icon("delta", mean_color), mean_diff_display, "Mean % Difference", mean_color,
      caption = "average % difference vs. what's published"
    ),
    kpi_tile(svg_icon("plus", added_color), length(added), "Added", added_color, caption = condense_changes(added)),
    kpi_tile(svg_icon("minus", removed_color), length(removed), "Removed", removed_color, caption = condense_changes(removed))
  )
})
```

## Row {height=160px}

```{r}
uiOutput("kpi_boxes_ui")
```

## Row

```{r}
#| title: Indicator Metadata
reactableOutput("summary_table")
```

# Completeness Check

*This section verifies whether the internal dataset covers the same country-year combinations as the public CEPALSTAT version, for the currently filtered countries and years.* It flags "Missing in Both", "Old Only" (removed in this update), "New Only" (added in this update), and "Present in Both".

```{r}
#| context: server
plot_coverage <- function(df) {
  df2 <- df %>%
    mutate(
      country = factor(country, levels = sort(unique(country), decreasing = TRUE)),
      status = factor(status, levels = c("Present in Both", "New Only", "Old Only", "Missing in Both"))
    )
  ggplot(df2, aes(
    x = year, y = country, color = status,
    text = paste0("Country: ", country, "<br>Year: ", year, "<br>Status: ", status)
  )) +
    geom_point(shape = 15, size = 3.2) +
    scale_color_manual(values = c(
      "Missing in Both" = pal$fill_missing,
      "Old Only"        = pal$before,
      "New Only"        = pal$after,
      "Present in Both" = pal$fill_both
    ), drop = FALSE) +
    labs(x = NULL, y = NULL, color = "Coverage") +
    theme_qc() +
    theme(legend.position = "bottom", panel.grid.major = element_line(color = pal$grid, linewidth = 0.2))
}

output$coverage_plot <- renderPlotly({
  fd <- filtered_data()
  ggplotly(plot_coverage(fd$coverage_check), tooltip = "text") %>%
    style_plotly()
})
```

```{r}
#| title: Data Availability by Country and Year
plotlyOutput("coverage_plot", height = "650px")
```

# Published Preview

*Previews how this update will change CEPALSTAT's own published time series graphic, for the Highlight section's selected Country (and dimension member(s), if the indicator has an extra dimension).* Each line pair shares one color: solid is the current internal value, muted dot-dash is the value currently published. If the dimension filter has more than one member selected (or is left at "(All)"), each member gets its own color and line pair - narrow it to one member to see a single line, matching CEPALSTAT's own chart exactly. The x-axis follows the sidebar's Year range.

```{r}
#| context: server
plot_published_preview <- function(agg_df, year_range) {
  members <- sort(unique(agg_df$member))
  n <- length(members)
  base_colors <- if (n <= length(pal$categorical)) {
    pal$categorical[seq_len(n)]
  } else {
    grDevices::hcl.colors(n, "Dark3")
  }
  names(base_colors) <- members

  long_df <- agg_df %>%
    filter(year >= year_range[1], year <= year_range[2]) %>%
    pivot_longer(cols = c(value_data, value_pub), names_to = "vintage", values_to = "value") %>%
    mutate(
      vintage = recode(vintage, value_data = "Current (internal)", value_pub = "Published"),
      series = paste0(member, " — ", vintage)
    ) %>%
    filter(!is.na(value))

  color_map <- character(0)
  linetype_map <- character(0)
  for (m in members) {
    color_map[paste0(m, " — Current (internal)")] <- base_colors[[m]]
    color_map[paste0(m, " — Published")] <- mute_color(base_colors[[m]])
    linetype_map[paste0(m, " — Current (internal)")] <- "solid"
    linetype_map[paste0(m, " — Published")] <- "dotdash"
  }

  p <- ggplot(long_df, aes(
    x = year, y = value, color = series, linetype = series, group = series,
    text = paste0("Year: ", year, "<br>", member, " (", vintage, ")", "<br>Value: ", round(value, 1))
  )) +
    geom_line(linewidth = 0.9) +
    scale_color_manual(values = color_map) +
    scale_linetype_manual(values = linetype_map) +
    scale_x_continuous(breaks = scales::pretty_breaks()) +
    scale_y_continuous(labels = label_number()) +
    labs(x = NULL, y = NULL, color = NULL, linetype = NULL) +
    theme_qc() +
    theme(legend.position = "bottom")

  p + geom_point(
    data = function(d) dplyr::filter(d, vintage == "Current (internal)"),
    size = 1.8, show.legend = FALSE
  )
}

output$published_preview_plot <- renderPlotly({
  d <- qc_data()
  if (is.null(input$country)) return(empty_plotly("Pick a country in the sidebar."))
  if (is.null(input$year_range)) return(empty_plotly("Loading..."))

  dim_sel <- if (d$has_extra_dims) {
    setNames(list(input[[paste0("dimfilter_", d$extra_dims[1])]]), d$extra_dims[1])
  } else {
    list()
  }
  agg <- build_published_preview(d, input$country, dim_sel)
  if (nrow(agg) == 0) return(empty_plotly("No data for this selection."))

  p <- plot_published_preview(agg, input$year_range)
  ggplotly(p, tooltip = "text") %>% style_plotly(legend_bottom = TRUE)
})
```

```{r}
#| title: Current vs. Published
plotlyOutput("published_preview_plot", height = "480px")
```

# New Values

This section visualizes the values of the new data points, both across every filtered country and for the country selected in the sidebar's Inspect Country dropdown.

```{r}
#| context: server
output$new_overview <- renderPlotly({
  fd <- filtered_data()
  df2 <- fd$comp %>% filter(!is.na(value_data))
  p <- ggplot(df2, aes(
    x = as.numeric(dim_29117_label), y = value_data, group = dim_208_label, color = flag_new_entry,
    text = paste0(
      "Country: ", dim_208_label, "<br>Year: ", dim_29117_label,
      "<br>Value: ", round(value_data, 1), "<br>New Entry: ", flag_new_entry
    )
  )) +
    geom_line(alpha = 0.35, color = pal$grid) +
    geom_point(size = 1.3, alpha = 0.85) +
    scale_color_manual(values = c("FALSE" = pal$neutral, "TRUE" = pal$after)) +
    labs(title = "Time Series by Country", x = "Year", y = "Value") +
    theme_qc() +
    theme(legend.position = "none")
  ggplotly(p, tooltip = "text") %>% style_plotly(legend_bottom = FALSE)
})

output$new_country <- renderPlotly({
  fd <- filtered_data()
  if (is.null(input$country)) return(empty_plotly("Pick a country in the sidebar."))
  df2 <- fd$comp %>% filter(!is.na(value_data), dim_208_label == input$country)
  p <- ggplot(df2, aes(
    x = as.numeric(dim_29117_label), y = value_data, color = flag_new_entry,
    text = paste0("Year: ", dim_29117_label, "<br>Value: ", round(value_data, 1), "<br>New Entry: ", flag_new_entry)
  )) +
    geom_line(color = pal$grid, linewidth = 0.7) +
    geom_point(size = 2.6, alpha = 0.9) +
    scale_color_manual(values = c("FALSE" = pal$neutral, "TRUE" = pal$after)) +
    labs(title = input$country, x = "Year", y = "Value") +
    theme_qc() +
    theme(legend.position = "none")
  ggplotly(p, tooltip = "text") %>% style_plotly(legend_bottom = FALSE)
})

# All-countries faceted view (kept alongside the reactive single-country view for now, as a
# comprehensive cross-check while the filter-driven approach gets a trial run).
output$new_faceted <- renderPlotly({
  fd <- filtered_data()
  df2 <- fd$comp %>% filter(!is.na(value_data))
  p <- ggplot(df2, aes(
    x = as.numeric(dim_29117_label), y = value_data, group = dim_208_label, color = flag_new_entry,
    text = paste0(
      "Country: ", dim_208_label, "<br>Year: ", dim_29117_label,
      "<br>Value: ", round(value_data, 1), "<br>New Entry: ", flag_new_entry
    )
  )) +
    geom_line(alpha = 0.4, color = pal$grid) +
    geom_point(size = 1, alpha = 0.85) +
    facet_wrap(~ dim_208_label, scales = "free_y", ncol = 3) +
    scale_color_manual(values = c("FALSE" = pal$neutral, "TRUE" = pal$after)) +
    labs(title = "All Filtered Countries", x = "Year", y = "Value") +
    theme_qc() +
    theme(legend.position = "none")
  ggplotly(p, tooltip = "text") %>% style_plotly(legend_bottom = FALSE)
})

output$new_faceted_ui <- renderUI({
  fd <- filtered_data()
  scrollable_plot("new_faceted", facet_height_px(n_distinct(fd$comp$dim_208_label)))
})
```

## Row {height=420px}

```{r}
#| title: Time Series by Country (All Filtered Countries)
plotlyOutput("new_overview", height = "420px")
```

```{r}
#| title: Inspect Country
plotlyOutput("new_country", height = "420px")
```

## Row {height=740px}

```{r}
#| title: All Countries (Faceted)
uiOutput("new_faceted_ui")
```

# Value Comparisons

*This section analyzes changes between values that are available in both the public and internal data sets, for the currently filtered countries/years.* Key statistics include mean and absolute differences, and the share of values with changes greater than ±20%. Red points identify values that exceed this threshold. Skipped automatically for a new indicator with no published data to compare against.

```{r}
#| context: server
no_pub_msg <- "This is a new indicator, so there is no previously published data to compare against."

output$summary_stats_table <- renderReactable({
  d <- qc_data()
  if (!d$has_public_data) return(reactable(tibble(Message = no_pub_msg)))
  fd <- filtered_data()
  stats <- fd$comp %>%
    filter(status == "Present in Both") %>%
    summarise(
      "Mean % difference" = round(mean(perc_diff[is.finite(perc_diff)], na.rm = TRUE), 1),
      "% of values with large differences (>20%)" = round(mean(flag_large_diff, na.rm = TRUE) * 100, 1),
      "Mean absolute difference" = round(mean(abs_diff[is.finite(abs_diff)], na.rm = TRUE), 1),
      "Median reported value" = round(median(value_data, na.rm = TRUE), 1),
      "Minimum reported value" = round(min(value_data, na.rm = TRUE), 1),
      "Maximum reported value" = round(max(value_data, na.rm = TRUE), 1)
    )
  df <- tibble(Metric = names(stats), Value = as.character(unlist(stats)))
  reactable(df, columns = list(Metric = colDef(width = 280)), bordered = TRUE, highlight = TRUE)
})

output$scatter_plot <- renderPlotly({
  d <- qc_data()
  if (!d$has_public_data) return(empty_plotly(no_pub_msg))
  fd <- filtered_data()
  df2 <- fd$comp %>%
    filter(!flag_some_na) %>%
    mutate(flag_large_diff = factor(flag_large_diff, levels = c(FALSE, TRUE), labels = c("Within ±20%", "Above ±20%")))
  p <- ggplot(df2, aes(
    x = value_data, y = value_pub, color = flag_large_diff,
    text = paste0(
      "Country: ", dim_208_label, "<br>Year: ", dim_29117_label,
      "<br>Internal: ", round(value_data, 1), "<br>Public: ", round(value_pub, 1)
    )
  )) +
    geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = pal$ink2) +
    geom_point(alpha = 0.8, size = 1.5) +
    scale_color_manual(values = c("Within ±20%" = pal$neutral, "Above ±20%" = pal$flag), guide = "none") +
    labs(title = "Value Comparison between Public and Internal Data", x = "Internal Data", y = "Public Data") +
    theme_qc()
  ggplotly(p, tooltip = "text")
})

output$pctdiff_plot <- renderPlotly({
  d <- qc_data()
  if (!d$has_public_data) return(empty_plotly(no_pub_msg))
  fd <- filtered_data()
  yr <- range(as.numeric(fd$comp$dim_29117_label), na.rm = TRUE)
  df2 <- fd$comp %>% filter(!flag_some_na)
  p <- ggplot(df2, aes(
    x = as.numeric(dim_29117_label), y = perc_diff,
    text = paste0("Country: ", dim_208_label, "<br>Year: ", dim_29117_label, "<br>% Diff: ", round(perc_diff, 1))
  )) +
    annotate("rect", xmin = yr[1] - 0.5, xmax = yr[2] + 0.5, ymin = -20, ymax = 20, fill = pal$good, alpha = 0.08) +
    geom_jitter(aes(color = abs(perc_diff) > 20), width = 0.2, height = 0, size = 1.4, alpha = 0.8) +
    scale_color_manual(values = c("FALSE" = pal$neutral, "TRUE" = pal$flag), guide = "none") +
    labs(title = "Percent Change between Public and Internal Data", x = "Year", y = "Percent Difference") +
    theme_qc()
  ggplotly(p, tooltip = "text") %>% style_plotly(legend_bottom = FALSE)
})

output$threshold_table <- renderReactable({
  d <- qc_data()
  if (!d$has_public_data) return(reactable(tibble(Message = no_pub_msg)))
  fd <- filtered_data()
  flagged <- fd$comp %>%
    mutate(
      z_perc_diff = round(as.numeric(scale(perc_diff)), 1),
      value_data = round(value_data, 1),
      value_pub = round(value_pub, 1),
      perc_diff = round(perc_diff, 1)
    ) %>%
    filter(flag_large_diff == TRUE) %>%
    arrange(desc(z_perc_diff)) %>%
    select(dim_208_label, dim_29117_label, value_data, value_pub, perc_diff, z_perc_diff)
  if (nrow(flagged) == 0) {
    return(reactable(tibble(Message = "No values differ from the published data by more than ±20%.")))
  }
  reactable(flagged, searchable = TRUE, bordered = TRUE, highlight = TRUE, defaultPageSize = 8)
})
```

```{r}
#| title: Summary Statistics
reactableOutput("summary_stats_table")
```

## Row {height=420px}

```{r}
#| title: Value Comparison (Public vs. Internal)
plotlyOutput("scatter_plot", height = "420px")
```

```{r}
#| title: Percent Change Over Time
plotlyOutput("pctdiff_plot", height = "420px")
```

```{r}
#| title: Records Above ±20% Difference
reactableOutput("threshold_table")
```

# Outlier Detection

*This section evaluates the quality of the internal data independently, flagging records where the value is more than 3 standard deviations from the country's own mean.* This works on the internal data alone, so it applies whether or not there is public data to compare against.

```{r}
#| context: server
output$outlier_table <- renderReactable({
  fd <- filtered_data()
  out <- fd$comp %>%
    filter(flag_some_na == FALSE, outlier_within_country == TRUE) %>%
    mutate(value_data = round(value_data, 1)) %>%
    arrange(desc(abs(z_within_country))) %>%
    select(dim_208_label, dim_29117_label, value_data, z_within_country)
  if (nrow(out) == 0) {
    return(reactable(tibble(Message = "No within-country outliers were found.")))
  }
  reactable(out, searchable = TRUE, bordered = TRUE, highlight = TRUE, defaultPageSize = 8)
})

output$outlier_country_plot <- renderPlotly({
  fd <- filtered_data()
  if (is.null(input$country)) return(empty_plotly("Pick a country in the sidebar."))
  df2 <- fd$comp %>% filter(dim_208_label == input$country)
  p <- ggplot(df2, aes(
    x = as.numeric(dim_29117_label), y = value_data, color = flag_some_value,
    text = paste0(
      "Year: ", dim_29117_label, "<br>Internal: ", round(value_data, 1),
      "<br>Public: ", round(value_pub, 1), "<br>% Diff: ", round(perc_diff, 1),
      "<br>Z-Score: ", round(z_within_country, 1)
    )
  )) +
    geom_line(color = pal$grid, linewidth = 0.7) +
    geom_point(size = 2.6, alpha = 0.9) +
    scale_color_manual(values = c("FALSE" = pal$neutral, "TRUE" = pal$flag)) +
    labs(title = input$country, x = "Year", y = "Internal Value") +
    theme_qc() +
    theme(legend.position = "none")
  ggplotly(p, tooltip = "text") %>% style_plotly(legend_bottom = FALSE)
})

# All-countries faceted view (kept alongside the reactive single-country view for now).
output$outlier_faceted <- renderPlotly({
  fd <- filtered_data()
  p <- ggplot(fd$comp, aes(
    x = as.numeric(dim_29117_label), y = value_data, group = dim_208_label, color = flag_some_value,
    text = paste0(
      "Country: ", dim_208_label, "<br>Year: ", dim_29117_label,
      "<br>Internal: ", round(value_data, 1), "<br>Public: ", round(value_pub, 1),
      "<br>% Diff: ", round(perc_diff, 1), "<br>Z-Score: ", round(z_within_country, 1)
    )
  )) +
    geom_line(alpha = 0.4, color = pal$grid) +
    geom_point(size = 1, alpha = 0.85) +
    facet_wrap(~ dim_208_label, scales = "free_y", ncol = 3) +
    scale_color_manual(values = c("FALSE" = pal$neutral, "TRUE" = pal$flag)) +
    labs(title = "All Filtered Countries", x = "Year", y = "Internal Value") +
    theme_qc() +
    theme(legend.position = "none")
  ggplotly(p, tooltip = "text") %>% style_plotly(legend_bottom = FALSE)
})

output$outlier_faceted_ui <- renderUI({
  fd <- filtered_data()
  scrollable_plot("outlier_faceted", facet_height_px(n_distinct(fd$comp$dim_208_label)))
})
```

## Row {height=420px}

```{r}
#| title: Z-Score Outliers (Filtered View)
reactableOutput("outlier_table")
```

```{r}
#| title: Inspect Country
plotlyOutput("outlier_country_plot", height = "420px")
```

## Row {height=740px}

```{r}
#| title: All Countries (Faceted)
uiOutput("outlier_faceted_ui")
```

# Dimension Analysis

*Breaks out the selected country's values by the indicator's extra dimension (e.g. sex or age group).* Always shows every member of that dimension, regardless of the dimension filter in the sidebar, since that's the point of this view. Only applies to indicators with an extra dimension beyond country and year - if there's ever more than one, only the first is used here.

```{r}
#| context: server
no_dims_msg <- "This indicator has no extra dimensions beyond country and year."

output$dim_plot <- renderPlotly({
  d <- qc_data()
  if (!d$has_extra_dims) return(empty_plotly(no_dims_msg))
  if (is.null(input$country)) return(empty_plotly("Pick a country in the sidebar."))
  dim_col <- d$extra_dims[1]
  dim_label <- d$dim_labels[[dim_col]]

  by_extra <- d$comp_detailed %>%
    filter(dim_208_label == input$country, !is.na(value_data)) %>%
    group_by(across(all_of(c("dim_29117_label", dim_col)))) %>%
    summarize(value_data = sum(value_data, na.rm = TRUE), .groups = "drop") %>%
    mutate(dim_29117_label = as.numeric(dim_29117_label))

  n_cats <- n_distinct(by_extra[[dim_col]])
  color_scale <- if (n_cats <= length(pal$categorical)) {
    scale_color_manual(values = pal$categorical[seq_len(n_cats)])
  } else {
    scale_color_viridis_d(option = "C")
  }

  p <- ggplot(by_extra, aes(
    x = dim_29117_label, y = value_data,
    color = .data[[dim_col]], group = .data[[dim_col]],
    text = paste0("Year: ", dim_29117_label, "<br>", dim_label, ": ", .data[[dim_col]], "<br>Value: ", round(value_data, 1))
  )) +
    geom_line(linewidth = 0.8) +
    geom_point(size = 1.8) +
    color_scale +
    scale_x_continuous(n.breaks = 6) +
    scale_y_continuous(labels = label_number()) +
    labs(title = glue("{input$country} — {dim_label}"), x = NULL, y = NULL, color = dim_label) +
    theme_qc() +
    theme(legend.position = "bottom")

  ggplotly(p, tooltip = "text") %>% style_plotly(legend_bottom = TRUE)
})

# All-countries faceted view (kept alongside the reactive single-country view for now).
output$dim_faceted <- renderPlotly({
  d <- qc_data()
  if (!d$has_extra_dims) return(empty_plotly(no_dims_msg))
  dim_col <- d$extra_dims[1]
  dim_label <- d$dim_labels[[dim_col]]

  by_extra <- d$comp_detailed %>%
    filter(!is.na(value_data)) %>%
    group_by(across(all_of(c("dim_208_label", "dim_29117_label", dim_col)))) %>%
    summarize(value_data = sum(value_data, na.rm = TRUE), .groups = "drop") %>%
    mutate(dim_29117_label = as.numeric(dim_29117_label))

  n_cats <- n_distinct(by_extra[[dim_col]])
  color_scale <- if (n_cats <= length(pal$categorical)) {
    scale_color_manual(values = pal$categorical[seq_len(n_cats)])
  } else {
    scale_color_viridis_d(option = "C")
  }

  p <- ggplot(by_extra, aes(
    x = dim_29117_label, y = value_data, color = .data[[dim_col]], group = .data[[dim_col]],
    text = paste0(
      "Country: ", dim_208_label, "<br>Year: ", dim_29117_label,
      "<br>", dim_label, ": ", .data[[dim_col]], "<br>Value: ", round(value_data, 1)
    )
  )) +
    geom_line(linewidth = 0.7) +
    facet_wrap(~ dim_208_label, scales = "free_y", ncol = 3) +
    color_scale +
    scale_x_continuous(n.breaks = 4) +
    scale_y_continuous(labels = label_number(), n.breaks = 4) +
    labs(title = glue("All Filtered Countries — {dim_label}"), x = NULL, y = NULL, color = dim_label) +
    theme_qc() +
    theme(legend.position = "bottom")

  ggplotly(p, tooltip = "text") %>% style_plotly(legend_bottom = TRUE)
})

output$dim_faceted_ui <- renderUI({
  d <- qc_data()
  req(d$has_extra_dims)
  fd <- filtered_data()
  scrollable_plot("dim_faceted", facet_height_px(n_distinct(fd$comp$dim_208_label)))
})
```

```{r}
#| title: Country x Dimension Breakout
plotlyOutput("dim_plot", height = "480px")
```

## Row {height=740px}

```{r}
#| title: All Countries (Faceted)
uiOutput("dim_faceted_ui")
```
