---
title: "Quality Check Report"
format: 
  html:
    self-contained: true
    toc: true
    toc_float: true
    code_folding: show
    theme: cosmo
    math: "mathjax"
params:
  indicator_id: 4049
execute:
  echo: false
  warning: false
  message: false
project:
  type: default
  output-dir: "QC Reports"
output-file: "qc_report_4049.html"
---

## Quality Check Report for Indicator `r params$indicator_id`

<!--
This report summarizes the quality checks performed to validate updated internal data for environmental indicators.
It compares internal data with the publicly available CEPALSTAT dataset to assess completeness, consistency, and statistical plausibility.
Non-technical readers are encouraged to read the section summaries for guidance on how to interpret results.
-->

```{r setup}
library(tidyverse)
library(magrittr)
library(readxl)
library(glue)
library(knitr)
library(gt)
library(reactable)
library(plotly)
library(scales)
library(lubridate)
library(DT)
library(scales)
library(here)
library(fs)

source(here("Scripts/utils.R"))

# Load data
indicator_id <- params$indicator_id
comp <- read_xlsx(here(glue("Data/Checks/comp_id{indicator_id}.xlsx")))

# Overwrite types
comp %<>%
  mutate(across(
    .cols = matches("dim"),
    .fns = as.character)) %>%
  mutate(across(
    .cols = matches("value|^abs_diff$|^perc_diff$"),
    .fns = as.numeric)) %>%
  mutate(across(
    .cols = starts_with("flag_"),
    .fns = as.logical))

iso <- read_xlsx(here("Data/iso_codes.xlsx"))

iso %<>% 
  filter(ECLACa == "Y") %>% 
  select(cepalstat, name, std_name)

meta <- read_xlsx(here("Data/indicator_metadata.xlsx"))
indicator_name <- meta %>% filter(id == indicator_id) %>% pull(indicator)

date_stamp <- Sys.Date()

# Get data vintage
files <- dir_ls(here("Data/Cleaned"), regexp = paste0("id", indicator_id, "_\\d{4}-\\d{2}-\\d{2}T\\d{6}"))

latest_file <- tibble(
  file = files,
  datetime = str_extract(files, "\\d{4}-\\d{2}-\\d{2}T\\d{6}") %>%
    ymd_hms(truncated = 1)) %>%
  arrange(desc(datetime)) %>%
  slice(1) %>%
  pull(file)

data_date <- str_extract(latest_file, "\\d{4}-\\d{2}-\\d{2}T\\d{6}")

# Get dimension info
ind_dims <- comp %>% select(ends_with("_label")) %>% names()
base_dims <- c("dim_29117_label", "dim_208_label") # year, country
extra_dims <- setdiff(ind_dims, base_dims)
extra_dim_ids <- str_extract(extra_dims, "\\d+")

extra_dim_names <- get_indicator_dimensions(indicator_id) %>% 
  filter(id %in% extra_dim_ids) %>% 
  pull(name)

dim_ids <- if (length(extra_dim_ids) > 0) paste(extra_dim_ids, collapse = ", ") else ""
dim_names <- if (length(extra_dim_names) > 0) paste(extra_dim_names, collapse = ", ") else ""

# Create summary info
summary_info <- list(
  "Report Date" = format(date_stamp, "%d %B %Y"),
  "Indicator ID" = indicator_id,
  "Indicator Name" = indicator_name,
  "Data Vintage" = data_date,
  "Data Source" = meta %>% filter(id == indicator_id) %>% pull(source),
  "Dimension IDs" = dim_ids,
  "Dimension Names" = dim_names,
  "Time Coverage" = paste(min(comp$dim_29117_label), "-", max(comp$dim_29117_label)),
  "Number of Countries" = n_distinct(comp$dim_208_label),
  "Total Records" = nrow(comp)
)
```

<!--
Create comparison dfs at two different levels of detail:
- comp is summarized to the country/year level and drops the record_id field
- comp_detailed is left in the original data format: includes record_id, dim_XXXX, dim_XXXX_label for any extra dims
-->

```{r}
# Keep `comp` df at country/year level. Create comp_detailed for IF there are additional dimensions.
comp_detailed <- comp

## Summarize comp to country/year level
comp %<>%
  group_by(dim_208, dim_29117, dim_208_label, dim_29117_label) %>%
  summarize(
    value_data = if (all(is.na(value_data))) NA_real_ else sum(value_data, na.rm = TRUE),
    value_pub  = if (all(is.na(value_pub)))  NA_real_ else sum(value_pub,  na.rm = TRUE),
    .groups = "drop"
  )

comp %<>% 
   mutate(abs_diff = abs(value_data - value_pub),
         perc_diff = round((abs_diff / value_pub) * 100, 2)) %>% 
  mutate(
    flag_large_diff = perc_diff > 20,  # adjust threshold as needed
    flag_missing_entry = is.na(value_data), # exists in pub but not in data
    flag_new_entry = is.na(value_pub),     # entry exists in data but not in pub
    flag_some_na = is.na(value_data) | is.na(value_pub) # either new or missing entry
  ) %>% 
  mutate(status = case_when(
    is.na(value_data) & is.na(value_pub) ~ "Missing in Both",
    is.na(value_data) ~ "Old Only",
    is.na(value_pub) ~ "New Only",
    TRUE ~ "Present in Both"
  ))

#setdiff(names(comp_detailed), names(comp))
```


## Summary Info

This section summarizes key metadata about the indicator and dataset under review.
It helps contextualize the scope of the data quality checks.

```{r}
# Create summary table
summary_df <- tibble(
  Metric = names(summary_info),
  Value = unlist(summary_info)
)

summary_df %>%
  gt() %>%
  tab_header(
    title = "QC Report Information",
    subtitle = glue("Generated on {format(date_stamp, '%d %B %Y')}")
  ) %>%
  tab_style(
    style = list(
      cell_fill(color = "#f7f7f7"),
      cell_text(weight = "bold")
    ),
    locations = cells_column_labels()
  )
```

## Completeness Check

### Data Completeness by Country and Year

*This section verifies whether the internal dataset covers the same country-year combinations as the public CEPALSTAT version.* 
It flags:

- "Missing in Both" (no data in either source)

- "Old Only" (data was removed or lost in the new version)

- "New Only" (new values added that were not in the original public version)

- "Present in Both" (overlap between datasets)

The visual below shows data coverage by country and year. Gaps or additions should be reviewed for justification.


```{r}
all_combinations <- expand_grid(
  country = unique(iso$std_name),
  year = as.character(min(comp$dim_29117_label):max(comp$dim_29117_label))
)

coverage_check <- all_combinations %>%
  left_join(comp, by = c("country" = "dim_208_label", "year" = "dim_29117_label")) %>% 
  mutate(status = ifelse(is.na(status), "Missing in Both", status))

# Remove sub-regions if not in data set ## CHANGE TO ASSOCIATES
regional_status <- coverage_check$status[which(coverage_check$country %in% c("Latin America", "Caribbean", "South America", "Central America"))]

if(all(regional_status == "Missing in Both")) {
  coverage_check %<>%
    filter(!country %in% c("Latin America", "Caribbean", "South America", "Central America"))
}

```

```{r}
p <- coverage_check %>% 
  mutate(
    country = factor(country, levels = sort(unique(country), decreasing = TRUE)),
    status = factor(status, 
                   levels = c("Present in Both", "New Only", "Old Only", "Missing in Both"),
                   labels = c("Present in Both", "New Only", "Old Only", "Missing in Both"))
  ) %>% 
  ggplot(aes(x = year, y = country, fill = status,
             text = paste("Country:", country, 
                         "<br>Year:", year, 
                         "<br>Status:", status))) +
  geom_tile(color = "white") +
  scale_fill_manual(values = c(
    "Missing in Both" = "white",
    "Old Only" = "#ff9999",
    "New Only" = "#99ccff",
    "Present in Both" = "#e6e6e6"
  )) +
  labs(
    title = "Data Availability by Country and Year",
    subtitle = "Shows data availability in both internal and public datasets",
    x = NULL, 
    y = NULL, 
    fill = "Data Coverage"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    plot.subtitle = element_text(size = 12),
    legend.position = "bottom",
    legend.title = element_text(face = "bold")
  )

ggplotly(p, tooltip = "text") %>%
  layout(
    legend = list(orientation = "h", y = -0.2)
  )
```


### New Values Time Series

This section visualizes the values of the new data points. Whether any of the new data points for the new year are anomalous can be explored in the Country-Level Analysis section.


```{r}
p <- ggplot(comp, aes(
  x = as.numeric(dim_29117_label),  # Year
  y = value_data,
  group = dim_208_label,            # Country
  color = flag_new_entry,
  text = paste(
    "Country:", dim_208_label,
    "<br>Year:", dim_29117_label,
    "<br>Value:", round(value_data, 1),
    "<br>New Entry:", flag_new_entry
  )
)) +
  # Base lines by country
  geom_line(alpha = 0.4, color = "#e6e6e6") +
  
  # Highlight new entries
  geom_point(size = 1.3, alpha = 0.8) +
  
  # Color scheme for new entries
  scale_color_manual(values = c("FALSE" = "gray80", "TRUE" = "#99ccff")) +
  
  # Labels and theme
  labs(
    title = "Time Series by Country",
    subtitle = "Red points indicate new entries not present in public dataset",
    x = "Year",
    y = "Value"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    plot.subtitle = element_text(size = 12),
    legend.position = "none"
  )

# Interactive conversion
ggplotly(p, tooltip = "text") %>%
  layout(
    title = list(
      text = paste0(
        "Time Series by Country",
        '<br><sup>Blue points indicate new entries not present in public dataset</sup>'
      )
    )
  )

```

### Country Values Time Series
This shows the same information but is scaled to the country level. 

```{r, warning = FALSE, fig.height=14}
p <- comp %>% 
  filter(!is.na(value_data)) %>% 
  ggplot(aes(
  x = as.numeric(dim_29117_label),  # Year
  y = value_data,
  group = dim_208_label,            # Country
  color = flag_new_entry,
  text = paste(
    "Country:", dim_208_label,
    "<br>Year:", dim_29117_label,
    "<br>Value:", round(value_data, 1),
    "<br>New Entry:", flag_new_entry
  )
)) +
  # Base lines by country
  geom_line(alpha = 0.5, color = "#e6e6e6") +
  
  # Graph by individual country
  facet_wrap(~ dim_208_label, scales = "free_y", ncol = 3) +
  
  # Highlight new entries
  geom_point(size = 1, alpha = 0.8) +
  
  # Color scheme for new entries
  scale_color_manual(values = c("FALSE" = "gray80", "TRUE" = "#99ccff")) +
  
  # Labels and theme
  labs(
    title = "Individual Country Time Series",
    subtitle = "Red points indicate new entries not present in public dataset",
    x = "Year",
    y = "Value"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    plot.subtitle = element_text(size = 12),
    legend.position = "none"
  )

# Interactive conversion
ggplotly(p, tooltip = "text") %>%
  layout(
    title = list(
      text = paste0(
        "Time Series by Country"#,
        #'<br><sup>Blue points indicate new entries not present in public dataset</sup>'
      )
    )
  )
```

## Value Comparisons (Between Data Sets)

*This section analyzes changes between values that are available in both the public and internal data sets.*
Key statistics include mean and absolute differences, and the share of values with changes greater than ±20%.
Red dots in the plots identify values that exceed this threshold.

This helps identify outliers, input errors, or significant data revisions that may need explanation.

### Summary Statistics

```{r}
summary_info <- comp %>%
  filter(status == "Present in Both") %>%
  summarise(
    "Mean % difference" = round(mean(perc_diff[is.finite(perc_diff)], na.rm = TRUE), 1),
    "% of values with large differences (<20%)" = round(mean(flag_large_diff, na.rm = TRUE) * 100, 1),
    "Mean absolute difference" = round(mean(abs_diff[is.finite(abs_diff)], na.rm = TRUE), 1),
    "Median reported value" = round(median(value_data, na.rm = TRUE), 1),
    "Minimum reported value" = round(min(value_data, na.rm = TRUE), 1),
    "Maximum reported value" = round(max(value_data, na.rm = TRUE), 1)
  )

# Convert to long format
summary_df <- tibble(
  Metric = names(summary_info),
  Value = unlist(summary_info)
)

# Format with % where applicable
summary_df <- summary_df %>%
  mutate(
    Value = case_when(
      grepl("%", Metric) ~ percent(as.numeric(Value) / 100, accuracy = 0.1),
      TRUE ~ as.character(Value)
    )
  )

# Render with gt
summary_df %>%
  gt() %>%
  tab_header(
    title = "Change Detection Summary",
    subtitle = "Key metrics comparing values present in both internal and public datasets"
  ) %>%
  tab_style(
    style = list(
      cell_fill(color = "#f7f7f7"),
      cell_text(weight = "bold")
    ),
    locations = cells_column_labels()
  )
```


### Absolute Value Comparison

This section compares data values for country/year pairs between the public and internal data sets. The dashed line indicates perfect alignment between the two sources. Any points colored in red indicate a greater than 20% difference between the two data sets.

```{r}
p <- comp %>% 
  filter(flag_some_na == FALSE) %>% # remove partial entries
  mutate(flag_large_diff = factor(flag_large_diff, 
                             levels = c(FALSE, TRUE),
                             labels = c("Within ±20%", "Above ±20%"))) %>% 
  ggplot(aes(
  y = value_pub,
  x = value_data,
  color = flag_large_diff,
  text = paste(
    "<br>Country:", dim_208_label,
    "<br>Year:", dim_29117_label,
    "<br>Public:", round(value_pub, 1),
    "<br>Internal:", round(value_data, 1),
    "<br>Above Threshold:", flag_large_diff
  )
)) +
  geom_point(alpha = 0.8, size = 1.4) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray60") +
  scale_color_manual(
    values = c("Within ±20%" = "#99ccff", "Above ±20%" = "#cc6666"),
    guide = "none"
  ) +
  labs(
    y = "Public Data",
    x = "Internal Data",
    title = "Value Comparison between Public and Internal Data",
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    plot.subtitle = element_text(size = 12),
    legend.position = "bottom",
    legend.title = element_text(face = "bold")
  )

# Convert to interactive plot with tooltips
ggplotly(p, tooltip = "text")
```

### Relative Difference Comparison

This section compares relative differences between the public and internal data. A percent change is calculated between the public data and internal data, and the data is displayed in country/year pairs. The ±20% tolerance threshold band is colored a light green. Large relative changes above 20% are colored red.

```{r}
year_min <- min(as.numeric(comp$dim_29117_label), na.rm = TRUE)
year_max <- max(as.numeric(comp$dim_29117_label), na.rm = TRUE)

p <- comp %>% 
  filter(flag_some_na == FALSE) %>% 
  ggplot(aes(
    x = as.numeric(dim_29117_label),
    y = perc_diff,
    text = paste(
      "<br>Country:", dim_208_label,
      "<br>Year:", dim_29117_label,
      "<br>% Diff:", round(perc_diff, 1)
    )
  )) +
  # Fix: Band fill outside aes()
  geom_rect(xmin = year_min, xmax = year_max, ymin = -20, ymax = 20,
            fill = "#e6f4ea", alpha = 0.6, inherit.aes = FALSE) +

  geom_jitter(aes(color = abs(perc_diff) > 20), 
              width = 0.2, height = 0, size = 1.4, alpha = 0.8) +

  scale_color_manual(values = c("FALSE" = "#99ccff", "TRUE" = "#cc6666"), guide = "none") +

  labs(
    title = "Percent Change between Public and Internal Data",
    subtitle = "Red points are outside of the ±20% tolerance threshold",
    x = "Year",
    y = "Percent Difference"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    plot.subtitle = element_text(size = 12),
    legend.position = "none"
  )

ggplotly(p, tooltip = "text") %>%
  layout(
    title = list(
      text = paste0(
        "Percent Change between Public and Internal Data",
        '<br><sup>Red points are outside of the ±20% tolerance threshold</sup>'
      )
    )
  )


```



### Percent Difference Threshold Table

This table displays records where the updated internal values differ from the public CEPALSTAT data by more than 20%, corresponding to the red points in the plots above.

Rows with a z-score (`z_perc_diff`) greater than 3 are statistically significant outliers, indicating values that deviate strongly from the overall distribution and may require further review or explanation.

```{r}
reactable(
  comp %>%
    mutate(
      z_perc_diff = as.numeric(scale(perc_diff)),  # force numeric vector
      value_data = round(value_data, 1),
      value_pub = round(value_pub, 1),
      perc_diff = round(perc_diff, 1),
      z_perc_diff = round(z_perc_diff, 1)
    ) %>%
    filter(flag_large_diff == TRUE) %>%
    arrange(desc(z_perc_diff)) %>%
    select(
      #record_id,
      dim_208_label,
      dim_29117_label,
      value_data,
      value_pub,
      perc_diff,
      z_perc_diff
    ),
  defaultPageSize = 5,
  searchable = TRUE,
  highlight = TRUE,
  bordered = TRUE,
  theme = reactableTheme(
    style = list(fontFamily = "Segoe UI, Roboto, Helvetica, sans-serif", fontSize = "14px")
  )
)

```

## Outlier Detection (Within Data Set)

*This section evaluates the quality of the internal data independently by providing statistical outlier detection based on z-scores.* It flags any records where the percent difference is more than 3 standard deviations from the mean, which could indicate outliers or extreme changes in the data.

A z-score measures how far a value is from the average, in terms of standard deviations. It’s calculated as: 

<div style="text-align:center">

**z = (x − μ) / σ**

</div>

where: x is the value of interest, μ is the mean of the group, and σ is the standard deviation.

Typical interpretations:

- A z-score of 0 means the value is at the mean

- A z-score of ±1 is within the typical range (68% of data is within 1 s.d.)

- A z-score of ±2 is unusual (95% of data is within 2 s.d.)

- A z-score of ±3 or more is considered an outlier (99.7% of data is within 3 s.d.)

<!--
Future additions:
Could add GLOBAL outlier detection that:
- Calculates global z-scores, though summary rows like LAC would have to be removed
- Could create checks to highlight any other potential structural issues such as implausible or negative values, or duplicates, or NAs
-->


### Country-Level Analysis

The following checks analyze the internal data within each country set. This helps differentiate any significant country-level anomalies that may be less apparent than at the global scale.

<!--
Future additions:
- Could create checks to highlight any sharp year-over-year changes by looking at % change from lag(year)
- Could create checks for any flat values with 0 standard deviation over time
-->

#### Z-Score Threshold Table

This calculates z-scores within each country series to flag major deviations in the individual country series. This displays all entries with a z-score greater than 3. If no entries are shown, then all of the values within each country seem reasonable.

```{r}
# Calculate and store country z-scores
comp %<>%
    group_by(dim_208_label) %>%
    mutate(
      z_within_country = round(as.numeric(scale(value_data)), 1),
      outlier_within_country = abs(z_within_country) > 3
    ) %>%
    ungroup()

# Create a flag for whether percent diff or z score alerts to a significant value diff
comp %<>% 
  mutate(flag_some_value = ifelse(outlier_within_country == TRUE | flag_large_diff == TRUE, TRUE, FALSE))
```


```{r}
reactable(
  comp %>%
    filter(flag_some_na == FALSE) %>%
    filter(outlier_within_country == TRUE) %>%
    mutate(
      value_data = round(value_data, 1),
      z_within_country = round(z_within_country, 1)
    ) %>%
    arrange(desc(abs(z_within_country))) %>%
    select(
      #record_id,
      dim_208_label,
      dim_29117_label,
      value_data,
      z_within_country
    ),
  defaultPageSize = 5,
  highlight = TRUE,
  bordered = TRUE,
  theme = reactableTheme(
    style = list(fontFamily = "Segoe UI, Roboto, Helvetica, sans-serif", fontSize = "14px")
  )
)

```


#### Outlier Country Plots

This section isolates country time series that have raised any flags: either with percent changes exceeding 20% or z-scores exceeding 3. This can be used to visually identify whether these data points are within-country outliers.

```{r}
# Step 1: Identify countries with any large difference or outlier
countries_with_flags <- comp %>% 
  filter(flag_some_na == FALSE & flag_some_value == TRUE) %>%
  distinct(dim_208_label) %>% 
  pull()

# Step 2: Styled interactive plot
p3 <- comp %>% 
  filter(dim_208_label %in% countries_with_flags) %>% 
  ggplot(aes(
    x = as.numeric(dim_29117_label),  # Year
    y = value_data,
    group = dim_208_label,
    color = flag_large_diff,
    text = paste(
      "Country:", dim_208_label,
      "<br>Year:", dim_29117_label,
      "<br>Internal Value:", round(value_data, 1),
      "<br>Public Value:", round(value_pub, 1),
      "<br>Public % Difference:", round(perc_diff, 1),
      "<br>Country Z-Score:", round(z_within_country, 1),
      "<br>Above % Threshold:", flag_large_diff
    )
  )) +
  # Subtle base lines for each country
  geom_line(alpha = 0.4, color = "gray70") +

  # Highlight flagged values
  geom_point(size = 1.2, alpha = 0.85) +

  # Color: red = flagged, blue = normal
  scale_color_manual(values = c("FALSE" = "#99ccff", "TRUE" = "#cc6666")) +

  labs(
    title = "Country Time Series with Flagged Values",
    subtitle = "Includes values flagged by percent difference or within-country z-score",
    x = "Year",
    y = "Internal Value"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    plot.subtitle = element_text(size = 12),
    legend.position = "none"
  )

# Convert to interactive plot
if(!is_empty(countries_with_flags)){
  ggplotly(p3, tooltip = "text") %>%
  layout(
    title = list(
      text = paste0(
        "Country Time Series with Flagged Values",
        '<br><sup>Red points indicate values flagged by percent difference or z-score</sup>'
      )
    )
  )
} else {
  plot(p3)
}


```

<!--
Need to add check for missing records when there are additional dimensions, or aggregate records
-->

```{r, warning = FALSE, fig.height=16, results='asis'}
render_plot_for_dim <- function(dim) {
  dim_name <- get_indicator_dimensions(indicator_id) %>% 
    filter(id == str_extract(dim, "\\d+")) %>% 
    pull(name)
  
  cat(glue::glue("### Dimension-Level Analysis"))
  
  by_extra <- comp_detailed %>%
    filter(!is.na(value_data)) %>% 
    group_by(across(all_of(c("dim_208_label", "dim_29117_label", dim)))) %>%
    summarize(value_data = sum(value_data, na.rm = TRUE),
              value_pub = sum(value_pub, na.rm = TRUE),
              .groups = "drop") %>%
    ungroup()

  p <- by_extra %>%
    mutate(dim_29117_label = as.numeric(dim_29117_label)) %>% 
    ggplot(aes(
      x = dim_29117_label,
      y = value_data,
      color = .data[[dim]],
      group = .data[[dim]],
      text = paste(
            "Country:", dim_208_label,
            "<br>Year:", dim_29117_label,
            "<br>Group:", .data[[dim]],
            "<br>Value:", round(value_data, 1)
          )
      
    )) +
    geom_line(linewidth = 0.8) +
    facet_wrap(~ dim_208_label, scales = "free_y", ncol = 3) +
    labs(
      title = glue("Country Breakout by Dimension: '{dim_name}'"),
      x = "Year",
      y = "",
      color = dim
    ) +
    scale_x_continuous(n.breaks = 4) +
    scale_y_continuous(labels = scales::label_number(), n.breaks = 4) +
    scale_color_viridis_d(option = "C") +
    theme_minimal() +
    theme(legend.position = "none")

  ggplotly(p, tooltip = "text")
}

if (length(extra_dims) >= 1) render_plot_for_dim(extra_dims[1])
if (length(extra_dims) == 2) render_plot_for_dim(extra_dims[2])
```

<!--
## Appendix

### Data Dictionary
-->
```{r}
# Create data dictionary
# data_dict <- tibble(
#   Field = names(comp),
#   Description = c(
#     "Unique record identifier",
#     "Country code",
#     "Country name",
#     "Year code",
#     "Year name",
#     "Value in internal dataset",
#     "Value in public dataset",
#     "Absolute difference between values",
#     "Percent difference between values",
#     "Flag for missing entries",
#     "Flag for new entries",
#     "Flag for large differences",
#     "Flag for NA values",
#     "Comparison status",
#     "Z score within country",
#     "Z score outlier within country",
#     "Flag for some value"
#   )
# )
# 
# data_dict %>%
#   gt() %>%
#   tab_header(
#     title = "Data Dictionary",
#     subtitle = "Description of fields in the comparison dataset"
#   )
```

<!--
### Technical Details
-->

<!--
Add code to spit out the name of the data being referenced. comp_idXXX currently doesn't store this. consider writing a second tab for the corresponding clean data file (since the comp files don't have time stamps)
-->

```{r}
# Session information
# sessionInfo()
```
